---
title: "Climate sensitivity"
format: docx
editor: visual
---

## Libraries and paths

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(stringdist)
library(writexl)
library(readr)
library(tidyr)
library(stringr)
library(broom)
library(ggplot2)
library(future.apply)
library(strucchange)
library(zoo)
library(patchwork)
library(WDI)
library(marginaleffects)
library(kableExtra)
library(lmtest)  
library(plm)
library(sp)
library(margins)
library(VIM)
library(ggbeeswarm)
#Some libararies are for testings purposes only


EUN <- c("AUT", "BEL", "BGR", "CYP", "CZE", "DEU", "DNK", "ESP", 
         "EST", "FIN", "FRA", "GRC", "HRV", "HUN", "IRL", "ITA", 
         "LTU", "LUX", "LVA", "MLT", "NLD", "POL", "PRT", "ROU", 
         "SVK", "SVN", "SWE")
e14 <- c("AUT", "BEL", "DNK", "FIN", "FRA", "DEU", "GRC", "IRL", 
         "ITA", "LUX", "NLD", "PRT", "ESP", "SWE")
nms <- c("BGR", "CYP", "CZE", "EST", "HRV", "HUN", "LVA", "LTU", 
         "MLT", "POL", "ROU", "SVK", "SVN")

# Define region groups:
ex_yug <- c("BIH", "HRV", "MKD", "MNE", "SRB", "SVN", "YUG") 
ex_ussr <- c("ARM", "AZE", "BLR", "EST", "GEO", "KAZ", "KGZ", 
             "LVA", "LTU", "MDA", "RUS", "TJK", "TKM", "UKR", "UZB")
former_sdn <- c("SDN", "SSDN")  
ETF   <- c("ERI","ETH")
BLX   <-c("BEL","LUX")
CSK   <-c("CZE","SVK")
CHN   <-c("HKG","MAC","CHN","TWN")
YSR   <-c("BIH","HRV","MKD","SVN","MNE","SRB","SER")
LANT  <-c("AIA","ATG","ABW","BRB","VGB","DMA","GRD","GLP","MTQ","MSR","ANT","KNA","LCA","VCT","VIR")
POLY  <-c("ASM","COK","PYF","NIU","WSM","TKL","TON","TUV","WLF")
MICRO <-c("GUM","KIR","MHL","FSM","NRU","MNP","PCI","PLW")
FIN   <- c("FIN","ALA")


datpth    <- "C:/Users/arsovsp/Documents/Climate Risk/data/"
datpthout <- "C:/Users/arsovsp/Documents/Climate Risk/output/"


map_wld <- read_rds(paste0(datpth,"globalmap.rds"))



```

#1.Load Yield Data

```{r}
#1. Load Yield Data
# Load raw FAO data and mapping files
country_mapping <- read.csv(paste0(datpth, "FAO_ISO_Country_List.csv"))
crop_mapping <- read.csv(paste0(datpth, "FAO_to_GAMS_Crop_Mapping.csv"))
fao_data <- read_excel(paste0(datpth, "FAOSTAT_CRP.xlsx"))

# Create mapping for element and crop names
element_mapping <- c("area harvested" = "AH", "yield" = "YLD", "production" = "QP")
crop_name_mapping <- c(
  "Barley" = "BA", "Oats" = "OT", "Rice" = "RI", 
  "Rape or colza seed" = "RP", "Rye" = "RY", 
  "Soya beans" = "SB", "Sunflower seed" = "SF", 
  "Sorghum" = "SO", "Wheat" = "WT", "Maize (corn)" = "MA"
)

# Preprocess FAO data
fao_data <- fao_data %>%
  mutate(
    Element = trimws(tolower(Element)),
    mmm = element_mapping[Element],
    ppp = crop_name_mapping[Item]
  ) %>%
  filter(!is.na(ppp) & !is.na(mmm))

# Rename and clean year columns
year_columns <- colnames(fao_data)[grepl("^Y", colnames(fao_data))]
colnames(fao_data)[grepl("^Y", colnames(fao_data))] <- sub("^Y", "", year_columns)
year_columns <- sub("^Y", "", year_columns)

# Convert units for yield
fao_data <- fao_data %>%
  mutate(across(all_of(year_columns),
                ~ ifelse(Element == "yield" & Unit == "100 g/ha", . / 10000, .))) %>%
  mutate(Unit = ifelse(Element == "yield" & Unit == "100 g/ha", "tons/ha", Unit))
country_mapping[which(country_mapping$ISO_Code== "TUR"),"FAO.Country.Name"]<-"Türkiye"
# Join ISO codes and filter
fao_data <- fao_data %>%
  left_join(country_mapping, by = c("Area" = "FAO.Country.Name")) %>%
  rename(ccc = ISO_Code) %>%
  filter(!is.na(ccc) & trimws(ccc) != "")

# Pivot to long format
fao_data_long <- fao_data %>%
  pivot_longer(
    cols = all_of(year_columns),
    names_to = "yyy",
    values_to = "value"
  ) %>%
  mutate(yyy = as.numeric(yyy))

# Pivot to wide: AH, YLD, QP
fao_data_wide <- fao_data_long %>%
  pivot_wider(
    id_cols = c(ccc, Item, ppp, yyy),
    names_from = mmm,
    values_from = value,
    values_fn = function(x) x[1],
    values_fill = list(value = NA)
  )

# Initial cleaned yield data
fao_yield_data <- fao_data_wide

# Add region mapping for all composites
fao_yield_data <- fao_yield_data %>%
  mutate(region_ccc = case_when(
    ccc %in% ex_yug ~ "YSR",
    ccc %in% ex_ussr ~ "USSR",
    ccc %in% former_sdn ~ "FSDN",
    ccc %in% BLX ~ "BLX",
    ccc %in% ETF ~ "ETF",
    TRUE ~ ccc
  ))

# --- STEP 1: FAO-supplied synthetic regions ---

fao_synthetic_regions <- fao_yield_data %>%
  filter(ccc %in% c("USSR", "YSR", "FSDN", "BLX", "ETF")) %>%
  filter(!is.na(YLD) & YLD > 0) %>%
  filter(!(ccc %in% c("USSR", "YSR") & yyy > 1991)) %>%  # cutoff only for USSR & YSR
  mutate(
    Yield_Imputed = FALSE,
    synthetic = TRUE,
    source = "FAO"
  )


# --- STEP 2: Aggregated yields from member countries ---
agg_yield_series <- fao_yield_data %>%
  filter(region_ccc %in% c("USSR", "YSR", "FSDN", "BLX", "ETF") & ccc != region_ccc) %>%
  group_by(region_ccc, yyy, ppp) %>%
  summarize(
    Total_Production = sum(QP, na.rm = TRUE),
    Total_Area = sum(AH, na.rm = TRUE),
    YLD = if_else(Total_Area == 0, NA_real_, Total_Production / Total_Area),
    AH = Total_Area,
    QP = Total_Production,
    Item = first(na.omit(Item)),
    .groups = "drop"
  ) %>%
  rename(ccc = region_ccc) %>%
  mutate(
    Yield_Imputed = TRUE,
    source = "Aggregated"
  )

# --- STEP 3: Clean join keys before removing overlaps ---
fao_synthetic_regions <- fao_synthetic_regions %>%
  mutate(across(c(ccc, ppp), as.character), yyy = as.numeric(yyy), ccc = trimws(ccc))
agg_yield_series <- agg_yield_series %>%
  mutate(across(c(ccc, ppp), as.character), yyy = as.numeric(yyy), ccc = trimws(ccc))

agg_yield_series_clean <- agg_yield_series %>%
  anti_join(fao_synthetic_regions %>% select(ccc, yyy, ppp), by = c("ccc", "yyy", "ppp"))

# --- STEP 4: Prepare Final dataset ---

fao_yield_data_clean <- fao_yield_data %>%
  filter(!ccc %in% c("USSR", "YSR", "FSDN", "BLX", "ETF"))

fao_yield_data_final <- bind_rows(
  fao_yield_data_clean,
  fao_synthetic_regions,
  agg_yield_series_clean
) %>%
  distinct(ccc, yyy, ppp, .keep_all = TRUE) %>%
  filter(!is.na(YLD) & YLD > 0)   # <<< REMOVE 0 or NA yields here



```

###1.1 Clean yield data

```{r}

# Plausible physical caps for yields (t/ha)
yld_caps <- tribble(
  ~ppp, ~yld_max,
  "MA", 15,   # Maize
  "BA",  9,   # Barley
  "WT",  9.5, # Wheat
  "RP",  5,   # Rapeseed
  "SB",  6.5, # Soybean
  "SF",  5    # Sunflower
)

min_yld <- 0.10        # treat <0.1 t/ha as invalid
rel_tol <- 0.10        # 10% tolerance vs QP/AH
max_spike_ratio <- 3   # spike if YLD_t / YLD_{t-1} > 3 or < 1/3


# =========================== 1) FLAGGING =========================
df_flagged <- fao_yield_data_final %>%
  left_join(yld_caps, by = "ppp") %>%
  arrange(ccc, ppp, yyy) %>%
  group_by(ccc, ppp) %>%
  mutate(
    # internal consistency
    yld_from_QP_AH = ifelse(!is.na(QP) & !is.na(AH) & AH > 0, QP / AH, NA_real_),
    flag_yld_mismatch = !is.na(yld_from_QP_AH) & !is.na(YLD) &
                        abs(YLD - yld_from_QP_AH) > rel_tol * yld_from_QP_AH,

    # plausibility
    flag_yld_low  = !is.na(YLD) & YLD < min_yld,
    flag_yld_high = !is.na(YLD) & !is.na(yld_max) & YLD > yld_max,

    # area/production sanity
    flag_AH_nonpos = !is.na(AH) & AH <= 0,
    flag_QP_nonpos = !is.na(QP) & QP <= 0,

    # spikes
    lag_yld  = dplyr::lag(YLD),
    ratio_up = ifelse(!is.na(lag_yld) & !is.na(YLD) & lag_yld > 0, YLD / lag_yld, NA_real_),
    flag_spike = !is.na(ratio_up) & (ratio_up > max_spike_ratio | ratio_up < 1 / max_spike_ratio),

    # series outliers (3*IQR fences)
    Q1  = quantile(YLD, 0.25, na.rm = TRUE),
    Q3  = quantile(YLD, 0.75, na.rm = TRUE),
    IQRv = Q3 - Q1,
    low_fence  = Q1 - 3 * IQRv,
    high_fence = Q3 + 3 * IQRv,
    flag_iqr_outlier = !is.na(YLD) & !is.na(low_fence) & !is.na(high_fence) &
                       (YLD < low_fence | YLD > high_fence)
  ) %>%
  ungroup() %>%
  mutate(
    flag_any = flag_yld_mismatch | flag_yld_low | flag_yld_high |
               flag_AH_nonpos   | flag_QP_nonpos | flag_spike | flag_iqr_outlier
  )

# Keep only your six crops
df_flagged_main <- df_flagged %>%
  filter(ppp %in% c("BA", "SB", "SF", "WT", "MA", "RP"))


# ============== 2) COUNTRY × CROP SUMMARY (flags) ===============
flag_summary_country_crop_main <- df_flagged_main %>%
  group_by(ccc, ppp) %>%
  summarise(
    n_obs            = n(),
    n_flag_any       = sum(flag_any, na.rm = TRUE),
    share_flag_any   = ifelse(n_obs > 0, n_flag_any / n_obs, NA_real_),
    n_mismatch       = sum(flag_yld_mismatch, na.rm = TRUE),
    n_low            = sum(flag_yld_low, na.rm = TRUE),
    n_high           = sum(flag_yld_high, na.rm = TRUE),
    n_spike          = sum(flag_spike, na.rm = TRUE),
    n_iqr_outlier    = sum(flag_iqr_outlier, na.rm = TRUE),
    n_AH_nonpos      = sum(flag_AH_nonpos, na.rm = TRUE),
    n_QP_nonpos      = sum(flag_QP_nonpos, na.rm = TRUE),
    .groups = "drop"
  )


# ======== 3) VERY PROBLEMATIC SERIES + SEVERITY SCORE ==========
very_problematic <- flag_summary_country_crop_main %>%
  mutate(
    hard_issue    = (n_high > 0) | (n_low >= 2) | (n_AH_nonpos > 0) | (n_QP_nonpos > 0),
    many_flags    = n_flag_any >= pmax(5, ceiling(0.15 * n_obs)),
    spiky         = n_spike >= 4,
    distro_out    = n_iqr_outlier >= 3,
    mismatch_many = n_mismatch >= 2,
    share_high    = (share_flag_any >= 0.15 & n_obs >= 20),

    very_problematic = hard_issue | many_flags | spiky | distro_out | mismatch_many | share_high,

    severity_score = 2.0*hard_issue +
                     1.5*share_high +
                     1.2*many_flags +
                     1.0*spiky +
                     0.8*distro_out +
                     0.5*mismatch_many +
                     0.3*(n_obs < 20)
  ) %>%
  filter(very_problematic) %>%
  arrange(desc(severity_score), desc(share_flag_any), desc(n_flag_any))

# (Optional) peek
# print(very_problematic)
# print(very_problematic %>% filter(severity_score >= 4))


# ===================== 4) DROP LISTS (with reasons) =====================
# A) Irrelevant / marginal country–crop pairs (manual list)
irrelevant_series <- tribble(
  ~ccc, ~ppp,
  "CPV","MA","QAT","MA","ARE","MA","KWT","MA",
  "OMN","BA","VCT","MA","TJK","RP","TJK","SB",
  "QAT","WT","ERI","MA","NAM","SF","SSD","SF","LSO","WT","CPV","MA"
) %>% mutate(reason = "irrelevant")

# B) Severity ≥ 4
problematic_high <- very_problematic %>%
  filter(severity_score >= 4) %>%
  transmute(ccc, ppp, reason = "severity_ge_4")

# C) Short AND noisy (fixing the earlier OR)
short_or_sparse <- flag_summary_country_crop_main %>%
  mutate(share_flag_any = n_flag_any / pmax(n_obs, 1)) %>%
  filter(n_obs < 20 & share_flag_any > 0.25) %>%      # AND: short AND noisy
  transmute(ccc, ppp, reason = "short_and_noisy")

# Combine unique pairs
to_drop <- bind_rows(
  irrelevant_series,
  problematic_high,
  short_or_sparse
) %>%
  distinct(ccc, ppp, .keep_all = TRUE)

# ===================== 5) KEY NORMALIZATION & AUDIT =====================
# Normalize join keys on BOTH sides (type/trim/case)
to_drop <- to_drop %>%
  mutate(across(c(ccc, ppp), ~ toupper(trimws(as.character(.)))))

fao_norm <- fao_yield_data_final %>%
  mutate(across(c(ccc, ppp), ~ toupper(trimws(as.character(.)))))

# Warn if drop list contains codes not present in data
not_in_data <- to_drop %>%
  anti_join(fao_norm %>% distinct(ccc, ppp), by = c("ccc","ppp"))
if (nrow(not_in_data)) {
  message("WARNING: entries in drop list not found in data (check codes):")
  print(not_in_data)
}

# ===================== 6) APPLY DROP + REPORT ===========================
fao_yield_data_final_clean <- fao_norm %>%
  anti_join(to_drop, by = c("ccc","ppp"))

# Exactly what was removed (and why)
removed_series <- fao_norm %>%
  semi_join(to_drop, by = c("ccc","ppp")) %>%
  distinct(ccc, ppp) %>%
  left_join(to_drop, by = c("ccc","ppp")) %>%
  arrange(reason, ppp, ccc)

cat("Dropped series:", nrow(removed_series), "\n")
cat("Remaining observations:", nrow(fao_yield_data_final_clean), "\n\n")

# Breakdown by reason
removed_series %>%
  count(reason, name = "n_series") %>%
  arrange(desc(n_series)) %>%
  print()

# Full list
print(removed_series, n = Inf)

#Winsorize under 0.1 ton of yields to 0.1
fao_yield_data_final_clean <- fao_yield_data_final_clean %>%
  mutate(YLD = ifelse(!is.na(YLD) & YLD < min_yld, min_yld, YLD))


# (Optional) replace original object after audit:
fao_yield_data_final <- fao_yield_data_final_clean


```

#2.Load and Reshape Climate Data

```{r}
# Get the list of countries from the FAO yield data (using original ISO codes)
fao_countries <- unique(fao_data$ccc)

# Define function to reshape climate data (assumes columns like "1950-01", "1950-02", …)
reshape_climate_data <- function(df, val_name) {
  df %>%
    pivot_longer(
      cols = matches("^[0-9]{4}-"),
      names_to = "YearMonth",
      values_to = val_name
    ) %>%
    mutate(
      yyy = as.numeric(sub("-.*", "", YearMonth)),
      Month = as.numeric(sub(".*-", "", YearMonth))
    ) %>%
    select(-YearMonth)
}

temp_mean <- read_excel(paste0(datpth, "era5-x0.25_timeseries_tas_timeseries_monthly_1950-2022_mean_historical_era5_x0.25_mean.xlsx")) %>%
  filter(code %in% fao_countries) %>% rename(ccc = code)
temp_min <- read_excel(paste0(datpth, "era5-x0.25_timeseries_tasmin_timeseries_monthly_1950-2022_mean_historical_era5_x0.25_mean.xlsx")) %>%
  filter(code %in% fao_countries) %>% rename(ccc = code)
temp_max <- read_excel(paste0(datpth, "era5-x0.25_timeseries_tasmax_timeseries_monthly_1950-2022_mean_historical_era5_x0.25_mean.xlsx")) %>%
  filter(code %in% fao_countries) %>% rename(ccc = code)
precip_data <- read_excel(paste0(datpth, "era5-x0.25_timeseries_pr_timeseries_monthly_1950-2022_mean_historical_era5_x0.25_mean.xlsx")) %>%
  filter(code %in% fao_countries) %>% rename(ccc = code)


temp_mean_long <- reshape_climate_data(temp_mean, "Temp_Mean")
temp_min_long  <- reshape_climate_data(temp_min, "Temp_Min")
temp_max_long  <- reshape_climate_data(temp_max, "Temp_Max")
precip_data_long <- reshape_climate_data(precip_data, "Precipitation")

temp_combined <- temp_mean_long %>%
  inner_join(temp_min_long, by = c("ccc", "yyy", "Month")) %>%
  inner_join(temp_max_long, by = c("ccc", "yyy", "Month"))



```

# 3. load and Fix ndvi data set

```{r}
# Step 1: Load FAOSTAT cropland file
fao_area_raw <- read_csv(paste0(datpth, "FAOSTAT_cropland.csv"))

# Build area weights from FAOSTAT cropland
area_data <- fao_area_raw %>%
  filter(
    Element == "Area",
    Item == "Cropland",
    Year == 2020
  ) %>%
  transmute(
    Area = Area,
    AreaValue = as.numeric(Value)
  ) %>%
  left_join(
    country_mapping %>%
      transmute(
        Area = `FAO.Country.Name`,
        iso3c = ISO_Code
      ),
    by = "Area"
  ) %>%
  filter(!is.na(iso3c)) %>%
  mutate(
    iso3c = ifelse(iso3c == "SSD", "SSDN", iso3c)
  ) %>%
  group_by(iso3c) %>%
  summarise(
    Area = sum(AreaValue, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Load NDVI dataset
ndvi_peak_bottom <- read_rds(paste0(datpth, "peak_bottom_ndvi_month_country.RDS"))
ndvi_season <- ndvi_peak_bottom %>% 
  select(ISO3, peak.cropland)

# Step 3: Aggregate for USSR & FSDN (BLX and ETF stay as-is in ndvi_season)
synthetic_ndvi <- ndvi_season %>%
  left_join(area_data, by = c("ISO3" = "iso3c")) %>%
  mutate(Area = ifelse(is.na(Area), 0, Area)) %>%
  filter(ISO3 %in% c(ex_ussr, former_sdn)) %>%
  mutate(region = case_when(
    ISO3 %in% ex_ussr ~ "USSR",
    ISO3 %in% former_sdn ~ "FSDN"
  )) %>%
  group_by(region) %>%
  summarize(
    peak.cropland = floor(weighted.mean(peak.cropland, Area, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  rename(ISO3 = region)

# Step 4: Append USSR & FSDN to the original set
ndvi_season <- bind_rows(ndvi_season, synthetic_ndvi)

print(ndvi_season %>% filter(ISO3 %in% c("BLX", "ETF", "USSR", "FSDN")))
summary(area_data)
```

###2.1. Compute Area-Weighted Climate Averages (Aggregated Regions)

```{r}

# Combine temperature and precipitation, and add area info
climate_data <- temp_combined %>%
  inner_join(precip_data_long, by = c("ccc", "yyy", "Month")) %>%
  left_join(area_data, by = c("ccc" = "iso3c")) %>%
  mutate(Area = ifelse(is.na(Area), 0, Area))  # <-- fill missing with 0 to avoid dropping rows

# Region mapping for composites
region_mapping <- data.frame(
  ccc = c(ex_yug, ex_ussr, former_sdn, BLX, ETF),
  region = c(rep("YSR", length(ex_yug)),
             rep("USSR", length(ex_ussr)),
             rep("FSDN", length(former_sdn)),
             rep("BLX", length(BLX)),
             rep("ETF", length(ETF))),
  stringsAsFactors = FALSE
)

# Tag each country with its composite region
climate_with_region <- climate_data %>%
  left_join(region_mapping, by = "ccc") %>%
  mutate(agg_ccc = if_else(is.na(region), ccc, region))

# Create synthetic monthly regional aggregates for USSR, YSR, FSDN, BLX, ETF
composite_climate <- climate_with_region %>%
  filter(!is.na(region)) %>%
  group_by(region, yyy, Month) %>%
  summarize(
    Temp_Mean = if (sum(Area, na.rm = TRUE) > 0) weighted.mean(Temp_Mean, Area, na.rm = TRUE) else mean(Temp_Mean, na.rm = TRUE),
    Temp_Max  = if (sum(Area, na.rm = TRUE) > 0) weighted.mean(Temp_Max, Area, na.rm = TRUE) else mean(Temp_Max, na.rm = TRUE),
    Temp_Min  = if (sum(Area, na.rm = TRUE) > 0) weighted.mean(Temp_Min, Area, na.rm = TRUE) else mean(Temp_Min, na.rm = TRUE),
    Precipitation = if (sum(Area, na.rm = TRUE) > 0) weighted.mean(Precipitation, Area, na.rm = TRUE) else mean(Precipitation, na.rm = TRUE),
    Area = sum(Area, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(ccc = region, agg_ccc = region)

# Combine back into the main dataset
climate_data_expanded <- bind_rows(climate_data, composite_climate)


```

### 2.2 Growing season based on ndvi

```{r}
# --- Merge NDVI into climate data
climate_seasonal <- climate_data_expanded %>%
  left_join(ndvi_season %>% select(ISO3, peak.cropland), by = c("ccc" = "ISO3"))

# --- Drop missing peak.cropland (these are non-represented composites)
climate_seasonal <- climate_seasonal %>%
  filter(!is.na(peak.cropland))

# --- Function for leap-year month lengths
get_month_lengths <- function(year) {
  is_leap <- (year %% 4 == 0 & (year %% 100 != 0 | year %% 400 == 0))
  days <- c(31, ifelse(is_leap, 29, 28), 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
  data.frame(Month = 1:12, Days = days)
}

l <- 5  # window length in months

# --- Build seasonal window per country-year based on peak.cropland
climate_seasonal <- climate_seasonal %>%
  group_by(ccc, yyy) %>%
  group_modify(~{
    center <- unique(.x$peak.cropland)
    if (is.na(center)) {
      # Return empty tibble with correct columns
      return(tibble(
        ccc = character(),
        yyy = numeric(),
        Month = numeric(),
        peak.cropland = numeric(),
        Temp_Mean = numeric(),
        Temp_Max = numeric(),
        Temp_Min = numeric(),
        Precipitation = numeric(),
        Area = numeric(),
        Days = numeric(),
        adj_year = numeric()
      ))
    }

    # Get leap-year-aware month lengths for this year
    month_len <- get_month_lengths(.y$yyy)
    .x <- .x %>% left_join(month_len, by = "Month")

    # Build centered window (wrap across year edges)
    season1 <- (center - (l - 1) / 2):(center + (l - 1) / 2)
    season2 <- ifelse(season1 <= 0, season1 + 12, season1)
    season2 <- ifelse(season2 > 12, season2 - 12, season2)

    d <- .x %>% filter(Month %in% season2)

    # Handle wrap-around (shift adj_year when season crosses Dec-Jan)
    base_year <- .y$yyy[1]
    if (length(unique(diff(season2))) != 1) {
      select <- season2[1:which(diff(season2) < 0)]
      d <- d %>% mutate(adj_year = if_else(Month %in% select, base_year + 1, base_year))
    } else {
      d <- d %>% mutate(adj_year = base_year)
    }

    d
  }) %>%
  ungroup()

```

###2.3 GDD Proxy using monthlies (Threshold of 5 C)

```{r}

threshold <- 5
gdd_seasonal <- climate_seasonal %>%
  group_by(ccc, season_year = adj_year) %>%   # use adj_year as the true "season year"
  summarize(
    Total_GDD = sum(pmax(((Temp_Max + Temp_Min) / 2) - threshold, 0) * Days, na.rm = TRUE),
    Avg_Temp_Mean = mean(Temp_Mean, na.rm = TRUE),
    Avg_Precip = mean(Precipitation, na.rm = TRUE),
    Tot_Precip = sum(Precipitation,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(ccc, season_year) %>%
  group_by(ccc) %>%
  mutate(GDD_Trend = Total_GDD - dplyr::lag(Total_GDD)) %>%
  ungroup() %>%
  rename(yyy = season_year)   # rename back so it merges with FAO



glimpse(gdd_seasonal)



```

###2.4. merge yield and clim Check for gaps

```{r}
# Merge the final yield dataset (fao_yield_data_final) with seasonal climate summary (gdd_seasonal)
yield_climate_data <- fao_yield_data_final %>%
  left_join(gdd_seasonal, by = c("ccc", "yyy")) %>%
  arrange(ccc, ppp, yyy)

glimpse(yield_climate_data)

# Define your threshold (e.g., remove series with > 1/3 bad values)
threshold <- 1/3

# Identify and keep only clean series
valid_series <- yield_climate_data %>%
  group_by(ccc, ppp) %>%
  summarize(
    n_total = n(),
    n_bad = sum(is.na(YLD) | YLD == 0),
    bad_ratio = n_bad / n_total,
    .groups = "drop"
  ) %>%
  filter(bad_ratio <= threshold)

# Filter main dataset to include only the good series
yield_climate_data_cleaned <- yield_climate_data %>%
  semi_join(valid_series, by = c("ccc", "ppp"))

# check how much is removed

n_removed <- yield_climate_data %>%
  anti_join(valid_series, by = c("ccc", "ppp")) %>%
  distinct(ccc, ppp) %>%
  nrow()

cat("Removed", n_removed, "country-crop pairs due to > 1/3 missing or zero YLD values.\n")

n_rows_removed <- yield_climate_data %>%
  anti_join(valid_series, by = c("ccc", "ppp")) %>%
  nrow()

cat("Total rows removed:", n_rows_removed)

#Total rows before
n_before <- yield_climate_data %>%
  distinct(ccc, ppp) %>%
  nrow()
print(n_before)

#Total rows after cleaning
n_after <- yield_climate_data_cleaned %>%
  distinct(ccc, ppp) %>%
  nrow()
print(n_after)


```

##3. Additional preparation of data (log, delta, square etc.)

###3.1 Prepare data for run (filter missing and so forth)

```{r}
## 1. Load Full FAO Data and Create a Wide Dataset
## 1. Load and Merge Data
selected_crops <- c("WT", "SF", "BA", "SB", "RP", "MA")
selected_countries <- NULL  

# Initial filtering
yield_filtered <- fao_yield_data_final %>%
  filter(ppp %in% selected_crops) %>%
  { if (!is.null(selected_countries)) filter(., ccc %in% selected_countries) else . }

climate_filtered <- gdd_seasonal %>%
  { if (!is.null(selected_countries)) filter(., ccc %in% selected_countries) else . }

# Merge and remove rows with missing core variables
yield_climate_data_cleaned <- yield_filtered %>% 
  left_join(climate_filtered, by = c("ccc", "yyy")) %>% 
  filter(!is.na(YLD), !is.na(Total_GDD), !is.na(Tot_Precip)) %>% 
  arrange(ccc, ppp, yyy)

## 2. Process and Filter (No Logs)
yield_climate_data_cleaned <- yield_climate_data_cleaned %>%
  group_by(ccc, ppp) %>%
  # Keep only series with 30+ years of data
  filter(n() >= 30) %>%
  mutate(
    # First differences of raw levels
    Delta_Yield  = YLD - dplyr::lag(YLD),
    Delta_GDD    = Total_GDD - dplyr::lag(Total_GDD),
    Delta_Precip = Tot_Precip - dplyr::lag(Tot_Precip)
  ) %>% 
  ungroup() %>%
  # Drop the first year of each group (where Delta is NA)
  filter(!is.na(Delta_Yield), !is.na(Delta_GDD), !is.na(Delta_Precip))

## 3. Clean Regions
# Ensure YSR and ex_ussr are defined as vectors before this
to_drop <- c(YSR, ex_ussr)

yield_climate_data_cleaned <- yield_climate_data_cleaned %>%
  filter(!ccc %in% to_drop)

# Final check
glimpse(yield_climate_data_cleaned)

```

###DELETE: 3.2 fix up of geo map data, and lat zones (informative only)

```{r}
# Identify the component countries of the USSR
ussr_countries <- c("RUS", "UKR", "BLR", "MDA", "GEO", "AZE", "ARM", "KAZ", "KGZ", "TJK", "TKM", "UZB")

# Aggregate the attributes of the component countries
ussr_data <- data.frame(
  FIPS = "US",
  ISO2 = "SU",
  ISO3 = "USSR",
  UN = mean(map_wld@data$UN[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  NAME = "USSR",
  AREA = sum(map_wld@data$AREA[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  POP2005 = sum(as.numeric(map_wld@data$POP2005[map_wld@data$ISO3 %in% ussr_countries]), na.rm = TRUE),
  REGION = mean(map_wld@data$REGION[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  SUBREGION = mean(map_wld@data$SUBREGION[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  LON = mean(map_wld@data$LON[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  LAT = mean(map_wld@data$LAT[map_wld@data$ISO3 %in% ussr_countries], na.rm = TRUE),
  REGION.NAME = "Asia",
  region = "Asia",
  FAO = "Rest of Europe and Central Asia"
)

# Create a new dataset that excludes the component countries of the USSR
new_map_wld_data <- map_wld@data[!map_wld@data$ISO3 %in% ussr_countries, ]

# Add the new row for the USSR to the new dataset
new_map_wld_data <- rbind(new_map_wld_data, ussr_data)

# Create a new SpatialPolygonsDataFrame with the merged USSR
new_map_wld <- map_wld
new_map_wld@data <- new_map_wld_data

#fixup sudam name
new_map_wld_data$ISO3[new_map_wld_data$ISO3 == "SDN"] <- "FSDN"
# However, a simpler version with 5 labels would be
# However, a simpler version with 5 labels would be
lat_zones <- cut(
  new_map_wld_data$LAT,
  breaks = c(-90, -35, -23.27, 23.27, 35, 90),
  labels = c("Temperate", "Sub‐tropical", "Tropical", "Sub‐tropical", "Temperate"),
  include.lowest = TRUE
)

# Add the latitude zones to your dataframe
new_map_wld_data$LAT_ZONE <- lat_zones
# Convert LAT_ZONE to character to avoid invalid factor level error
new_map_wld_data$LAT_ZONE <- as.character(new_map_wld_data$LAT_ZONE)

# Manually assign LAT_ZONE values
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "ETF"] <- "Tropical"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "CHN"] <- "Temperate"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "YSR"] <- "Temperate"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "LANT"] <- "Tropical"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "POLY"] <- "Tropical"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "MICRO"] <- "Temperate"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "BLX"] <- "Temperate"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "CSK"] <- "Temperate"
new_map_wld_data$LAT_ZONE[new_map_wld_data$ISO3 == "FIN"] <- "Polar"

# Convert LAT_ZONE back to factor if needed
new_map_wld_data$LAT_ZONE <- as.factor(new_map_wld_data$LAT_ZONE)
new_map_wld_data$LAT_ZONE <- as.character(new_map_wld_data$LAT_ZONE)
# Map the LAT_ZONE values to yield_climate_data
yield_climate_data_cleaned <- left_join(yield_climate_data_cleaned, new_map_wld_data[, c("ISO3", "LAT_ZONE")], by = c("ccc" = "ISO3"))

```

\*\*\*\*\*\*\*\*\*\*\*YIELD CLIMATE DATA CLEANED\*\*\*\*\*\*\*\*\*\* The data has beeen prepared up to this point and is located in

#4. MAIN FUNCTION

```{r}
run_break_analysis <- function(data,
                               selected_crops,
                               selected_countries = NULL,
                               model_formula,
                               window_years = 15,
                               global_test = FALSE) {

  mod_formula <- stats::as.formula(model_formula)

  # ---- templates (so bind_rows never breaks) ----
  tmpl_breaks <- tibble::tibble(country=character(), crop=character(), break_year=numeric(),
                                F_stat=numeric(), p_value=numeric(), significant=logical())
  tmpl_global <- tibble::tibble(country=character(), crop=character(),
                                global_p_value=numeric(), supF_stat=numeric(), supF_year=numeric(),
                                global_significant=logical())
  tmpl_mtn <- tibble::tibble(country=character(), crop=character(), yyy=numeric(), f_stat=numeric())
  tmpl_bic <- tibble::tibble(country=character(), crop=character(), m_breaks=integer(),
                             BIC_val=numeric(), is_chosen=logical())
  tmpl_vars <- tibble::tibble(country=character(), crop=character(), break_year=numeric(),
                              variable=character(), estimate_change=numeric(),
                              p_value=numeric(), significant=logical())
  tmpl_seg <- tibble::tibble(country=character(), crop=character(), segment=integer(),
                             start=numeric(), end=numeric(), r_squared=numeric(), coefficients=list())

  all_break_rows            <- list(tmpl_breaks)
  all_global_test_rows      <- list(tmpl_global)
  all_diagnostic_rows       <- list(tmpl_mtn)
  all_model_fit_rows        <- list(tmpl_bic)
  all_variable_details_rows <- list(tmpl_vars)
  all_segment_rows          <- list(tmpl_seg)
  results_list              <- list()

  # ---- filter ----
  filtered <- data %>%
    dplyr::filter(ppp %in% selected_crops) %>%
    { if (!is.null(selected_countries)) dplyr::filter(., ccc %in% selected_countries) else . } %>%
    dplyr::arrange(ccc, ppp, yyy) %>%
    tidyr::drop_na(dplyr::all_of(all.vars(mod_formula)))

  group_keys <- filtered %>% dplyr::distinct(ccc, ppp)

  for (i in seq_len(nrow(group_keys))) {

    ccc_i <- as.character(group_keys$ccc[i])
    ppp_i <- as.character(group_keys$ppp[i])

    ts_df <- filtered %>%
      dplyr::filter(ccc == ccc_i, ppp == ppp_i) %>%
      dplyr::arrange(yyy)

    n_obs <- nrow(ts_df)
    if (n_obs < (2 * window_years + 1)) next

    # ------------------------------------------------------------
    # 0) F-stats mountain + global test (length-synced alignment)
    # ------------------------------------------------------------
    skip_pair <- FALSE
    global_pval <- NA_real_
    fs <- tryCatch(
      strucchange::Fstats(mod_formula, data = ts_df,
                          from = window_years / n_obs,
                          to   = 1 - (window_years / n_obs)),
      error = function(e) NULL
    )

    if (!is.null(fs)) {
      f_vals <- as.numeric(fs$Fstats)
      L <- length(f_vals)

      if (L > 0) {
        start_pt <- floor((n_obs - L) / 2) + 1
        yrs <- ts_df$yyy[start_pt:(start_pt + L - 1)]

        all_diagnostic_rows[[length(all_diagnostic_rows) + 1]] <- tibble::tibble(
          country = ccc_i, crop = ppp_i, yyy = as.numeric(yrs), f_stat = as.numeric(f_vals)
        )

        global_pval <- tryCatch(strucchange::sctest(fs)$p.value, error = function(e) NA_real_)
        peak_idx <- which.max(f_vals)

        all_global_test_rows[[length(all_global_test_rows) + 1]] <- tibble::tibble(
          country = ccc_i, crop = ppp_i,
          global_p_value = as.numeric(global_pval),
          supF_stat = as.numeric(f_vals[peak_idx]),
          supF_year = as.numeric(yrs[peak_idx]),
          global_significant = isTRUE(!is.na(global_pval) && global_pval <= 0.05)
        )

        if (isTRUE(global_test) && (is.na(global_pval) || global_pval > 0.05)) skip_pair <- TRUE
      } else {
        if (isTRUE(global_test)) skip_pair <- TRUE
      }
    } else {
      if (isTRUE(global_test)) skip_pair <- TRUE
    }

    if (skip_pair) next

    # ------------------------------------------------------------
    # 1) Bai–Perron + full BIC path
    # ------------------------------------------------------------
    bp_full <- tryCatch(
      strucchange::breakpoints(mod_formula, data = ts_df, h = window_years),
      error = function(e) NULL
    )
    if (is.null(bp_full)) next

    bic_path <- tryCatch(stats::AIC(bp_full, k = log(n_obs)), error = function(e) NULL)

    bp_opt <- tryCatch(strucchange::breakpoints(bp_full), error = function(e) bp_full)
    opt_bps <- bp_opt$breakpoints
    n_breaks <- if (is.null(opt_bps) || all(is.na(opt_bps))) 0L else length(opt_bps)

    if (!is.null(bic_path)) {
      m_vec <- as.integer(seq_along(bic_path) - 1L)
      all_model_fit_rows[[length(all_model_fit_rows) + 1]] <- tibble::tibble(
        country = ccc_i, crop = ppp_i,
        m_breaks = m_vec,
        BIC_val = as.numeric(bic_path),
        is_chosen = (m_vec == n_breaks)
      )
    }

    bp_years <- if (n_breaks > 0) as.numeric(ts_df$yyy[opt_bps]) else numeric(0)

    # ------------------------------------------------------------
    # 2) Chow-style interaction test per breakpoint + shifts table
    #    (proper joint test: restricted vs unrestricted)
    # ------------------------------------------------------------
    this_pair_breaks <- list()
    this_pair_vars   <- list()

    if (n_breaks > 0) {

      # fit restricted once per pair
      mod_restricted <- tryCatch(stats::lm(mod_formula, data = ts_df), error = function(e) NULL)
      if (is.null(mod_restricted)) next

      for (byear in bp_years) {

        chow_df <- ts_df %>%
          dplyr::mutate(post = dplyr::if_else(yyy > byear, 1L, 0L))

        mod_unrestricted <- tryCatch(
          stats::lm(stats::update(mod_formula, . ~ . * post), data = chow_df),
          error = function(e) NULL
        )
        if (is.null(mod_unrestricted)) next

        # Joint F-test: are all post-related terms jointly zero?
        a12 <- tryCatch(stats::anova(mod_restricted, mod_unrestricted), error = function(e) NULL)
        if (is.null(a12) || nrow(a12) < 2) next

        fstat <- as.numeric(a12$F[2])
        pval  <- as.numeric(a12$`Pr(>F)`[2])

        row_b <- tibble::tibble(
          country = ccc_i, crop = ppp_i, break_year = as.numeric(byear),
          F_stat = fstat, p_value = pval,
          significant = isTRUE(!is.na(pval) && pval <= 0.05)
        )

        all_break_rows[[length(all_break_rows) + 1]] <- row_b
        this_pair_breaks[[length(this_pair_breaks) + 1]] <- row_b

        # Variable-level shifts (interaction coefficients + intercept shift)
        coef_out <- summary(mod_unrestricted)$coefficients
        rn <- rownames(coef_out)

        # "post" (intercept shift) and "something:post" (slope shifts)
        post_terms <- rn[grepl("(^post$)|(:post$)|(^post:)", rn)]

        for (term in post_terms) {
          var_l <- dplyr::case_when(
            term == "post" ~ "Intercept_Shift",
            TRUE ~ gsub("^post:|:post$", "", term)
          )

          row_v <- tibble::tibble(
            country = ccc_i, crop = ppp_i, break_year = as.numeric(byear),
            variable = var_l,
            estimate_change = as.numeric(coef_out[term, "Estimate"]),
            p_value = as.numeric(coef_out[term, "Pr(>|t|)"]),
            significant = isTRUE(as.numeric(coef_out[term, "Pr(>|t|)"]) <= 0.05)
          )

          all_variable_details_rows[[length(all_variable_details_rows) + 1]] <- row_v
          this_pair_vars[[length(this_pair_vars) + 1]] <- row_v
        }
      }
    }

    # ------------------------------------------------------------
    # 3) Segment models (non-overlapping, (lower, upper] intervals)
    # ------------------------------------------------------------
    this_pair_segments <- list()
    seg_bounds <- c(min(ts_df$yyy) - 1, sort(bp_years), max(ts_df$yyy))

    for (s in seq_len(length(seg_bounds) - 1)) {
      seg_df <- ts_df %>% dplyr::filter(yyy > seg_bounds[s], yyy <= seg_bounds[s + 1])
      if (nrow(seg_df) < window_years) next

      seg_mod <- tryCatch(stats::lm(mod_formula, data = seg_df), error = function(e) NULL)
      if (is.null(seg_mod)) next

      row_s <- tibble::tibble(
        country = ccc_i, crop = ppp_i, segment = as.integer(s),
        start = seg_bounds[s] + 1,
        end   = seg_bounds[s + 1],
        r_squared = as.numeric(summary(seg_mod)$r.squared),
        coefficients = list(stats::coef(seg_mod))
      )

      all_segment_rows[[length(all_segment_rows) + 1]] <- row_s
      this_pair_segments[[length(this_pair_segments) + 1]] <- row_s
    }

    results_list[[paste(ccc_i, ppp_i, sep = "_")]] <- list(
      break_years = bp_years,
      chow_tests = dplyr::bind_rows(this_pair_breaks),
      segments = dplyr::bind_rows(this_pair_segments),
      variable_details = dplyr::bind_rows(this_pair_vars)
    )
  }

  breaks_table           <- dplyr::bind_rows(all_break_rows)
  global_test_table      <- dplyr::bind_rows(all_global_test_rows)
  reviewer_evidence      <- dplyr::bind_rows(all_diagnostic_rows)
  model_fit_table        <- dplyr::bind_rows(all_model_fit_rows)
  variable_breaks_table  <- dplyr::bind_rows(all_variable_details_rows)
  segment_table          <- dplyr::bind_rows(all_segment_rows)

  # Summary table: dominant break = max F_stat (per pair), plus global test info
  breaks_summary <- if (nrow(breaks_table) > 0) {
    breaks_table %>%
      dplyr::group_by(country, crop) %>%
      dplyr::summarise(
        break_years = paste(sort(unique(break_year)), collapse = ", "),
        any_significant = any(significant, na.rm = TRUE),
        max_F_stat = max(F_stat, na.rm = TRUE),
        year_max_F = break_year[which.max(F_stat)],
        chow_p_value = p_value[which.max(F_stat)],
        .groups = "drop"
      ) %>%
      dplyr::left_join(global_test_table, by = c("country", "crop"))
  } else {
    global_test_table
  }

  list(
    results = results_list,
    breaks_summary = breaks_summary,
    breaks_table = breaks_table,
    variable_breaks_table = variable_breaks_table,
    model_fit_table = model_fit_table,
    reviewer_evidence = reviewer_evidence,
    global_test_table = global_test_table,
    segment_table = segment_table
  )
}

```

#5.Run MODELS

```{r}
# Run the function on a quadratic form
break_output <- run_break_analysis(
  data = yield_climate_data_cleaned,
  selected_crops = c("WT", "BA", "MA", "SB", "RP","SF"),
  selected_countries = NULL,
  model_formula = "YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy",
  window_years = 15,
  global_test = FALSE
)


# Access your specific variable p-values
var_stats <- break_output$variable_breaks_table
```

# II. Chapter 2 define breaks, review, and graph

### 1. Select the break years and flag in main data set

Filter to one final break per crop–country.Only significant breaks (p ≤ 0.05).Pick the highest F‑stat in that set.

```{r}
#flagging all the breaks
final_breaks <- break_output$breaks_table %>%
  filter(p_value <= 0.05) %>%
  group_by(country, crop) %>%
  slice_max(F_stat, n = 1, with_ties = FALSE) %>%
  ungroup()

dominant_breaks <- final_breaks %>%
  transmute(country, crop, break_year)

#Update the data set
Dataset_wbreaks <- yield_climate_data_cleaned %>%
  inner_join(final_breaks %>% select(country, crop, break_year),
             by = c("ccc" = "country", "ppp" = "crop")) %>%
  mutate(flag_before_after = if_else(yyy < break_year, 0, 1))

Dataset_breaks <- yield_climate_data_cleaned %>%
  left_join(final_breaks %>% select(country, crop, break_year),
             by = c("ccc" = "country", "ppp" = "crop")) %>%
  mutate(flag_before_after = if_else(yyy < break_year, 0, 1))
#Flag goes to 0 if no break available
Dataset_breaks[which(is.na(Dataset_breaks$break_year)), "flag_before_after" ]<-0
Dataset_breaks[which(is.na(Dataset_breaks$break_year)), "break_year" ]<-2022
```

### 2. Fit lines on pre-after breaks

We now re‑fit models separately for before/after break periods on the fulll dataset with breaks years selected.

```{r}
segment_models <-Dataset_breaks %>%
  group_by(ccc, ppp, flag_before_after) %>%
  group_modify(~ {
    mod <- lm(YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy, data = .x)
    tibble(
      r_squared = summary(mod)$r.squared,
      adj_r_squared = summary(mod)$adj.r.squared,
      coefficients = list(coef(mod)),
      marginal_effects = list(avg_slopes(mod))
    )
  }) %>%
  ungroup()

```

### 3.Save the data

```{r}
#a) Full flagged dataset:
Dataset_breaks[which((Dataset_breaks$break_year)=="2022"), "flag_before_after" ]<-1
write.csv(Dataset_breaks,
          file.path(datpthout, paste0("Processed_data_breaks_flagged_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

# b) Segment models:

segment_models_flat <- segment_models %>%
  mutate(
    coefficients = map_chr(coefficients, ~ paste(names(.x), round(.x, 3), collapse = "; ")),
    marginal_effects = map_chr(marginal_effects, ~ paste0(.x$term, "=", round(.x$estimate, 4), collapse = "; "))
  )

write.csv(segment_models_flat,
          file.path(datpthout, paste0("Fitted_lines_breaks", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)


#a) Full flagged dataset:
write.csv(Dataset_wbreaks,
          file.path(datpthout, paste0("Processed_only_breaks_flagged_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

```

###4.1 Model diagnostics and model outputs

```{r}
# 1. Prepare the list with a safer selector
list_of_datasets <- list(
  "A_Model_Selection"   = break_output$breaks_summary,
  "B_Break_Drivers"     = break_output$variable_breaks_table,
  "C_Phase_Performance" = break_output$segment_table %>%
                           dplyr::select(-dplyr::any_of(c("coefficients", "coefs"))),
  "D_FStat_Mountains"   = break_output$reviewer_evidence,
  "E_BIC_Elbow_Paths"   = break_output$model_fit_table
)

# 2. Export to Excel
openxlsx::write.xlsx(list_of_datasets, 
                     file = file.path(datpthout, "Structural_Change_Appendices_FINAL.xlsx"), 
                     overwrite = TRUE)

message("Success! All tabs (including the F-Stat Mountains) have been exported.")
```

###4.2 Select example to demonstrate

```{r}
# ==========================================================================
# FRANCE CASE STUDY (PUBLICATION): DIAGNOSTIC FIGURE + SEPARATE AME TABLE
# - (a) F-stat paths + chosen break year
# - (b) BIC elbow + chosen m
# - (c1/c2) Proper AMEs (quadratic-aware) Before vs After, with stars for SHIFT
# - Separate publication table for crops with validated breaks
# ==========================================================================

# --- Packages ---
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(broom)
library(flextable)

recode <- dplyr::recode

stopifnot(exists("break_output"))
stopifnot(exists("yield_climate_data_cleaned"))
stopifnot(exists("datpthout"))

# --- 1) Settings ---
target_country <- "FRA"
france_crops   <- c("BA", "MA", "RP", "WT", "SB")

case_cols <- c(
  "BA" = "#D55E00",
  "MA" = "#0072B2",
  "RP" = "#009E73",
  "WT" = "#CC79A7",
  "SB" = "#E69F00"
)

base_formula <- YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy

pstars <- function(p) dplyr::case_when(
  is.na(p) ~ "",
  p < 0.01 ~ "***",
  p < 0.05 ~ "**",
  p < 0.10 ~ "*",
  TRUE ~ ""
)

# number formatter: scientific notation for tiny values
fmt_num <- function(x, digits = 3, tiny = 1e-3, sci_digits = 2) {
  if (length(x) == 0 || is.na(x)) return("—")
  if (abs(x) < .Machine$double.eps) return("0")
  if (abs(x) < tiny) return(formatC(x, format = "e", digits = sci_digits))
  sprintf(paste0("%.", digits, "f"), x)
}

coef_either <- function(model, terms) {
  cf <- stats::coef(model)
  terms <- as.character(terms)
  hit <- intersect(terms, names(cf))
  if (length(hit) == 0) return(0)
  unname(cf[hit[1]])
}

fmt_coef_from_model <- function(model, term_name, digits = 3, tiny = 1e-3, sci_digits = 2) {
  td <- broom::tidy(model)
  row <- td %>% filter(term == term_name)
  if (nrow(row) == 0) return("—")
  paste0(fmt_num(row$estimate[1], digits = digits, tiny = tiny, sci_digits = sci_digits), pstars(row$p.value[1]))
}

ame_gdd_simple <- function(model, df) {
  b1 <- coef_either(model, "Total_GDD")
  b2 <- coef_either(model, "I(Total_GDD^2)")
  mean(b1 + 2 * b2 * df$Total_GDD, na.rm = TRUE)
}

ame_prec_simple <- function(model, df) {
  b1 <- coef_either(model, "Tot_Precip")
  b2 <- coef_either(model, "I(Tot_Precip^2)")
  mean(b1 + 2 * b2 * df$Tot_Precip, na.rm = TRUE)
}

# --- 2) Dominant break per crop (highest F among significant Chow tests) ---
fr_breaks <- break_output$breaks_table %>%
  filter(country == target_country, crop %in% france_crops, significant) %>%
  group_by(crop) %>%
  slice_max(F_stat, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(crop, break_year)

fr_breaks <- tibble(crop = france_crops) %>%
  left_join(fr_breaks, by = "crop") %>%
  mutate(has_break = !is.na(break_year))

# --- 3) Panel (a): F-stat mountains ---
mountain_plot_data <- break_output$reviewer_evidence %>%
  filter(country == target_country, crop %in% france_crops) %>%
  left_join(fr_breaks, by = "crop") %>%
  mutate(crop = factor(crop, levels = france_crops))

mountain_labels <- mountain_plot_data %>%
  filter(!is.na(break_year)) %>%
  group_by(crop) %>%
  summarise(
    break_year = first(break_year),
    xmin = min(yyy, na.rm = TRUE),
    xmax = max(yyy, na.rm = TRUE),
    ymax = max(f_stat, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    place_left = break_year > (xmin + 0.75 * (xmax - xmin)),
    x_lab = if_else(place_left, break_year - 0.6, break_year + 0.6),
    hjust_lab = if_else(place_left, 1, 0),
    y_lab = ymax * 0.94
  )

p_mountain <- ggplot(mountain_plot_data, aes(x = yyy, y = f_stat, color = crop)) +
  geom_line(linewidth = 0.8) +
  geom_vline(
    aes(xintercept = break_year),
    linetype = "dashed",
    linewidth = 0.6,
    alpha = 0.85,
    na.rm = TRUE
  ) +
  geom_text(
    data = mountain_labels,
    aes(x = x_lab, y = y_lab, label = as.integer(break_year), hjust = hjust_lab),
    size = 3,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ crop, nrow = 1, scales = "free_y") +
  scale_color_manual(values = case_cols) +
  scale_x_continuous(expand = expansion(mult = c(0.03, 0.08))) +
  labs(
    title = "a) Structural break evidence (F-statistic paths)",
    x = NULL,
    y = "F-statistic"
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "grey95", color = NA),
    panel.grid.minor = element_blank()
  )

# --- 4) Panel (b): BIC elbow ---
bic_summary <- break_output$model_fit_table %>%
  filter(country == target_country, crop %in% france_crops) %>%
  mutate(crop = factor(crop, levels = france_crops))

p_elbow <- ggplot(bic_summary, aes(x = m_breaks, y = BIC_val, group = crop, color = crop)) +
  geom_line(linewidth = 0.7) +
  geom_point(
    aes(fill = is_chosen),
    shape = 21,
    size = 2.4,
    stroke = 0.4,
    color = "black"
  ) +
  scale_fill_manual(values = c("TRUE" = "black", "FALSE" = "white"), guide = "none") +
  facet_wrap(~ crop, nrow = 1, scales = "free_y") +
  scale_color_manual(values = case_cols) +
  labs(
    title = "b) Model selection (BIC over number of breaks)",
    x = "Number of breaks (m)",
    y = "BIC"
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "grey95", color = NA),
    panel.grid.minor = element_blank()
  )

# --- 5) Proper AMEs (quadratic-aware) + shift tests ---
fr_ts <- yield_climate_data_cleaned %>%
  filter(ccc == target_country, ppp %in% france_crops) %>%
  transmute(
    ppp        = as.character(ppp),
    yyy        = as.numeric(yyy),
    YLD        = as.numeric(YLD),
    Total_GDD  = as.numeric(Total_GDD),
    Tot_Precip = as.numeric(Tot_Precip)
  ) %>%
  filter(complete.cases(.)) %>%
  as.data.frame()

compute_ame_and_shift <- function(crop_code, byear) {
  df <- fr_ts %>%
    filter(ppp == crop_code) %>%
    arrange(yyy)

  if (nrow(df) < 10) {
    return(tibble(
      Crop = crop_code,
      Break = NA_integer_,
      has_break = FALSE,
      ame_gdd_before  = NA_real_,
      ame_gdd_after   = NA_real_,
      p_shift_gdd     = NA_real_,
      ame_prec_before = NA_real_,
      ame_prec_after  = NA_real_,
      p_shift_prec    = NA_real_
    ))
  }

  if (is.na(byear)) {
    mod0 <- lm(base_formula, data = df)

    return(tibble(
      Crop = crop_code,
      Break = NA_integer_,
      has_break = FALSE,
      ame_gdd_before  = ame_gdd_simple(mod0, df),
      ame_gdd_after   = NA_real_,
      p_shift_gdd     = NA_real_,
      ame_prec_before = ame_prec_simple(mod0, df),
      ame_prec_after  = NA_real_,
      p_shift_prec    = NA_real_
    ))
  }

  df <- df %>%
    mutate(post = as.integer(yyy >= byear))

  df_before <- df %>% filter(post == 0)
  df_after  <- df %>% filter(post == 1)

  if (nrow(df_before) < 5 || nrow(df_after) < 5) {
    return(tibble(
      Crop = crop_code,
      Break = as.integer(byear),
      has_break = FALSE,
      ame_gdd_before  = NA_real_,
      ame_gdd_after   = NA_real_,
      p_shift_gdd     = NA_real_,
      ame_prec_before = NA_real_,
      ame_prec_after  = NA_real_,
      p_shift_prec    = NA_real_
    ))
  }

  mod_full <- lm(update(base_formula, . ~ . * post), data = df)

  b_gdd_1 <- coef_either(mod_full, "Total_GDD")
  b_gdd_2 <- coef_either(mod_full, "I(Total_GDD^2)")
  d_gdd_1 <- coef_either(mod_full, c("Total_GDD:post", "post:Total_GDD"))
  d_gdd_2 <- coef_either(mod_full, c("I(Total_GDD^2):post", "post:I(Total_GDD^2)"))

  b_pr_1 <- coef_either(mod_full, "Tot_Precip")
  b_pr_2 <- coef_either(mod_full, "I(Tot_Precip^2)")
  d_pr_1 <- coef_either(mod_full, c("Tot_Precip:post", "post:Tot_Precip"))
  d_pr_2 <- coef_either(mod_full, c("I(Tot_Precip^2):post", "post:I(Tot_Precip^2)"))

  ame_gdd_before <- mean(b_gdd_1 + 2 * b_gdd_2 * df_before$Total_GDD, na.rm = TRUE)
  ame_gdd_after  <- mean((b_gdd_1 + d_gdd_1) + 2 * (b_gdd_2 + d_gdd_2) * df_after$Total_GDD, na.rm = TRUE)

  ame_prec_before <- mean(b_pr_1 + 2 * b_pr_2 * df_before$Tot_Precip, na.rm = TRUE)
  ame_prec_after  <- mean((b_pr_1 + d_pr_1) + 2 * (b_pr_2 + d_pr_2) * df_after$Tot_Precip, na.rm = TRUE)

  mod_nogddshift <- lm(
    YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy +
      post + post:Tot_Precip + post:I(Tot_Precip^2) + post:yyy,
    data = df
  )
  p_shift_gdd <- tryCatch(
    anova(mod_nogddshift, mod_full)$`Pr(>F)`[2],
    error = function(e) NA_real_
  )

  mod_noprecshift <- lm(
    YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy +
      post + post:Total_GDD + post:I(Total_GDD^2) + post:yyy,
    data = df
  )
  p_shift_prec <- tryCatch(
    anova(mod_noprecshift, mod_full)$`Pr(>F)`[2],
    error = function(e) NA_real_
  )

  tibble(
    Crop = crop_code,
    Break = as.integer(byear),
    has_break = TRUE,
    ame_gdd_before  = ame_gdd_before,
    ame_gdd_after   = ame_gdd_after,
    p_shift_gdd     = p_shift_gdd,
    ame_prec_before = ame_prec_before,
    ame_prec_after  = ame_prec_after,
    p_shift_prec    = p_shift_prec
  )
}

fr_ame <- fr_breaks %>%
  mutate(out = map2(crop, break_year, compute_ame_and_shift)) %>%
  select(out) %>%
  unnest(out) %>%
  mutate(
    Crop = factor(Crop, levels = france_crops),
    star_gdd  = pstars(p_shift_gdd),
    star_prec = pstars(p_shift_prec)
  )

fr_ame_plot <- fr_ame %>% filter(has_break)

to_long_for_panel <- function(df, var_before, var_after, star_col, label) {
  df %>%
    transmute(
      Crop,
      x_before = .data[[var_before]],
      x_after  = .data[[var_after]],
      stars    = .data[[star_col]]
    ) %>%
    mutate(variable = label) %>%
    pivot_longer(cols = c(x_before, x_after), names_to = "period_raw", values_to = "x") %>%
    mutate(Period = if_else(period_raw == "x_before", "Before", "After")) %>%
    select(Crop, variable, Period, x, stars)
}

ame_long <- bind_rows(
  to_long_for_panel(fr_ame_plot, "ame_gdd_before",  "ame_gdd_after",  "star_gdd",  "GDD"),
  to_long_for_panel(fr_ame_plot, "ame_prec_before", "ame_prec_after", "star_prec", "Precip")
) %>%
  mutate(Period = factor(Period, levels = c("Before", "After")))

ame_seg <- fr_ame_plot %>%
  transmute(
    Crop,
    gdd_x0 = ame_gdd_before, gdd_x1 = ame_gdd_after, gdd_star = star_gdd,
    prc_x0 = ame_prec_before, prc_x1 = ame_prec_after, prc_star = star_prec
  )

p_c1 <- ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey35") +
  geom_segment(
    data = ame_seg,
    aes(x = gdd_x0, xend = gdd_x1, y = Crop, yend = Crop),
    linewidth = 0.7,
    color = "grey60"
  ) +
  geom_point(
    data = ame_long %>% filter(variable == "GDD"),
    aes(x = x, y = Crop, color = Crop, fill = Period),
    shape = 21,
    size = 3.2,
    stroke = 0.7
  ) +
  geom_text(
    data = fr_ame_plot,
    aes(x = ame_gdd_after, y = Crop, label = star_gdd),
    nudge_y = 0.15,
    size = 3.4,
    color = "black"
  ) +
  scale_color_manual(values = case_cols) +
  scale_fill_manual(values = c("Before" = "white", "After" = "grey20"), guide = "none") +
  labs(
    title = "c1) GDD sensitivity (Before vs After)",
    x = "Average marginal effect of GDD",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

p_c2 <- ggplot() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey35") +
  geom_segment(
    data = ame_seg,
    aes(x = prc_x0, xend = prc_x1, y = Crop, yend = Crop),
    linewidth = 0.7,
    color = "grey60"
  ) +
  geom_point(
    data = ame_long %>% filter(variable == "Precip"),
    aes(x = x, y = Crop, color = Crop, fill = Period),
    shape = 21,
    size = 3.2,
    stroke = 0.7
  ) +
  geom_text(
    data = fr_ame_plot,
    aes(x = ame_prec_after, y = Crop, label = star_prec),
    nudge_y = 0.15,
    size = 3.4,
    color = "black"
  ) +
  scale_color_manual(values = case_cols) +
  scale_fill_manual(values = c("Before" = "white", "After" = "grey20"), guide = "none") +
  labs(
    title = "c2) Precip sensitivity (Before vs After)",
    x = "Average marginal effect of precipitation",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

p_c <- p_c1 + p_c2 + plot_layout(widths = c(1, 1))

final_plot <- (p_mountain / p_elbow / p_c) +
  plot_layout(heights = c(1.05, 1.0, 0.95)) +
  plot_annotation(
    title = "France: Structural break diagnostics (evidence, selection, and before/after effects)",
    subtitle = "Stars indicate significance of the regime shift in sensitivity (nested F-test on interaction terms)."
  )

print(final_plot)

ggsave(
  filename = file.path(datpthout, "France_StructuralBreak_Diagnostics.png"),
  plot = final_plot,
  width = 190,
  height = 120,
  units = "mm",
  dpi = 600,
  bg = "white"
)

ggsave(
  filename = file.path(datpthout, "France_StructuralBreak_Diagnostics.pdf"),
  plot = final_plot,
  width = 190,
  height = 120,
  units = "mm",
  device = cairo_pdf
)

# ==========================================================================
# PUBLICATION TABLE: crops with validated breaks
# ==========================================================================

fit_segment_table <- function(crop_code, byear) {
  df <- fr_ts %>%
    filter(ppp == crop_code) %>%
    arrange(yyy)

  if (is.na(byear)) return(NULL)

  df_before <- df %>% filter(yyy < byear)
  df_after  <- df %>% filter(yyy >= byear)

  if (nrow(df_before) < 10 || nrow(df_after) < 10) return(NULL)

  mod_before <- lm(base_formula, data = df_before)
  mod_after  <- lm(base_formula, data = df_after)

  tibble(
    crop = crop_code,
    term = c(
      "R²",
      "(Intercept)",
      "Time (yyy)",
      "Total GDD",
      "Total GDD²",
      "AME GDD",
      "Tot Precip",
      "Tot Precip²",
      "AME Precip"
    ),
    Before = c(
      fmt_num(summary(mod_before)$r.squared, digits = 3),
      fmt_coef_from_model(mod_before, "(Intercept)", digits = 3),
      fmt_coef_from_model(mod_before, "yyy", digits = 3),
      fmt_coef_from_model(mod_before, "Total_GDD", digits = 3),
      fmt_coef_from_model(mod_before, "I(Total_GDD^2)", digits = 3, tiny = 1e-3, sci_digits = 2),
      fmt_num(ame_gdd_simple(mod_before, df_before), digits = 4, tiny = 1e-3, sci_digits = 2),
      fmt_coef_from_model(mod_before, "Tot_Precip", digits = 3),
      fmt_coef_from_model(mod_before, "I(Tot_Precip^2)", digits = 3, tiny = 1e-3, sci_digits = 2),
      fmt_num(ame_prec_simple(mod_before, df_before), digits = 4, tiny = 1e-3, sci_digits = 2)
    ),
    After = c(
      fmt_num(summary(mod_after)$r.squared, digits = 3),
      fmt_coef_from_model(mod_after, "(Intercept)", digits = 3),
      fmt_coef_from_model(mod_after, "yyy", digits = 3),
      fmt_coef_from_model(mod_after, "Total_GDD", digits = 3),
      fmt_coef_from_model(mod_after, "I(Total_GDD^2)", digits = 3, tiny = 1e-3, sci_digits = 2),
      fmt_num(ame_gdd_simple(mod_after, df_after), digits = 4, tiny = 1e-3, sci_digits = 2),
      fmt_coef_from_model(mod_after, "Tot_Precip", digits = 3),
      fmt_coef_from_model(mod_after, "I(Tot_Precip^2)", digits = 3, tiny = 1e-3, sci_digits = 2),
      fmt_num(ame_prec_simple(mod_after, df_after), digits = 4, tiny = 1e-3, sci_digits = 2)
    )
  ) %>%
    pivot_longer(cols = c(Before, After), names_to = "Period", values_to = "value")
}

tab_long <- fr_breaks %>%
  filter(has_break) %>%
  mutate(out = map2(crop, break_year, fit_segment_table)) %>%
  filter(!map_lgl(out, is.null)) %>%
  select(out) %>%
  unnest(out)

pub_crops <- unique(tab_long$crop)

tab_wide <- tab_long %>%
  mutate(
    Period = factor(Period, levels = c("Before", "After")),
    col = paste0(crop, "_", Period)
  ) %>%
  select(term, col, value) %>%
  pivot_wider(names_from = col, values_from = value) %>%
  mutate(order = match(
    term,
    c("R²", "(Intercept)", "Time (yyy)", "Total GDD", "Total GDD²", "AME GDD",
      "Tot Precip", "Tot Precip²", "AME Precip")
  )) %>%
  arrange(order) %>%
  select(-order)

hdr2_map <- c(term = "")
for (cr in pub_crops) {
  hdr2_map[paste0(cr, "_Before")] <- "Before"
  hdr2_map[paste0(cr, "_After")]  <- "After"
}

group_titles <- c("", paste0("France ", pub_crops))
group_widths <- c(1, rep(2, length(pub_crops)))

ft_fr <- flextable(tab_wide) %>%
  set_header_labels(values = hdr2_map) %>%
  add_header_row(values = group_titles, colwidths = group_widths) %>%
  theme_booktabs() %>%
  align(align = "center", part = "all") %>%
  align(j = 1, align = "left", part = "all") %>%
  italic(i = ~ grepl("AME", term), j = 1) %>%
  bold(part = "header") %>%
  add_footer_lines(
    "Note: * p<0.1; ** p<0.05; *** p<0.01. Small-magnitude coefficients are reported in scientific notation when rounding would display them as zero."
  ) %>%
  autofit()

ft_fr

save_as_docx(
  ft_fr,
  path = file.path(datpthout, "France_Break_Table.docx")
)
```

#III. DESCRIPTIVE TABLES AND GRAPHS FOR

##1. Input data summary

```{r}
# Calculate series length per crop-country pair
series_lengths <- yield_climate_data_cleaned %>%
  group_by(ppp, ccc) %>%
  summarise(series_length = n_distinct(yyy), .groups = "drop")

# Summary stats per crop
series_summary <- series_lengths %>%
  group_by(ppp) %>%
  summarise(
    avg_length = mean(series_length),
    min_length = min(series_length),
    max_length = max(series_length),
    .groups = "drop"
  )

print(series_summary)

# Histogram of series length per crop
ggplot(series_lengths, aes(x = series_length)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
  facet_wrap(~ppp, scales = "free_y") +
  labs(
    title = "Distribution of Data Series Lengths per Crop",
    x = "Number of Years in Series",
    y = "Count of Country-Series"
  ) +
  theme_minimal()
```

#2. climate and yield stats

```{r}


# Required libraries

library(e1071)     # for skewness()

# 1. Series‐length per crop–country and histogram
panel_lengths <- yield_climate_data_cleaned %>%
  count(ppp, ccc, name = "n_years")

ggplot(panel_lengths, aes(n_years)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
  facet_wrap(~ppp, scales = "free_y") +
  labs(
    title = "Distribution of Data Series Lengths per Crop",
    x = "Number of Years in Series",
    y = "Count of Country–Series"
  ) +
  theme_minimal()

# 2. Panel‐length summary by crop
crop_summary <- panel_lengths %>%
  group_by(ppp) %>%
  summarise(
    n_panels   = n(),
    avg_years  = round(mean(n_years), 2),
    min_years  = min(n_years),
    max_years  = max(n_years),
    .groups = "drop"
  ) %>%
  rename(
    Crop                   = ppp,
    `Panels (n)`           = n_panels,
    `Avg years`            = avg_years,
    `Min years`            = min_years,
    `Max years`            = max_years
  )

kable(crop_summary, caption = "Panel‐Length Summary by Crop") %>%
  kable_styling(full_width = FALSE)

# 3. Descriptive stats of yield by crop
desc_yield <- yield_climate_data_cleaned %>%
  group_by(ppp) %>%
  summarise(
    n        = n(),
    Min      = round(min(YLD, na.rm = TRUE), 2),
    Mean     = round(mean(YLD, na.rm = TRUE), 2),
    `25th`   = round(quantile(YLD, 0.25, na.rm = TRUE), 2),
    Median   = round(median(YLD, na.rm = TRUE), 2),
    `75th`   = round(quantile(YLD, 0.75, na.rm = TRUE), 2),
    Max      = round(max(YLD, na.rm = TRUE), 2),
    SD       = round(sd(YLD, na.rm = TRUE), 2),
    IQR      = round(IQR(YLD, na.rm = TRUE), 2),
    CV       = round(SD / Mean, 2),
    Skewness = round(skewness(YLD, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  rename(Crop = ppp)

kable(desc_yield, caption = "Descriptive Statistics of Yield by Crop") %>%
  kable_styling(full_width = FALSE)

# 4. Climate stats (across all panels)
climate_stats <- yield_climate_data_cleaned %>%
  summarise(
    across(
      .cols = c(Total_GDD, Tot_Precip),
      .fns  = list(Min = min, Mean = mean, Max = max, SD = sd),
      .names = "{col}_{fn}"
    )
  ) %>%
  mutate(across(everything(), ~round(.x, 2)))

kable(climate_stats, caption = "Aggregate Climate Statistics") %>%
  kable_styling(full_width = FALSE)

# 5. Country–Crop yield + outlier report
country_crop_report <- yield_climate_data_cleaned %>%
  group_by(ccc, ppp) %>%
  summarise(
    yld_min      = min(YLD, na.rm = TRUE),
    yld_mean     = mean(YLD, na.rm = TRUE),
    yld_max      = max(YLD, na.rm = TRUE),
    yld_sd       = sd(YLD, na.rm = TRUE),
    Q1           = quantile(YLD, 0.25, na.rm = TRUE),
    Q3           = quantile(YLD, 0.75, na.rm = TRUE),
    IQR          = IQR(YLD, na.rm = TRUE),
    lower_bound  = Q1 - 1.5 * IQR,
    upper_bound  = Q3 + 1.5 * IQR,
    outlier_obs  = sum(YLD < lower_bound | YLD > upper_bound, na.rm = TRUE),
    total_obs    = n(),
    pct_outliers = round(100 * outlier_obs / total_obs, 2),
    .groups      = "drop"
  ) %>%
  select(ccc, ppp, yld_min:pct_outliers)

# Export
write.csv(
  country_crop_report,
  file = file.path(datpthout, "country_crop_yield_report.csv"),
  row.names = FALSE
)



```

#3. Sptaial distribution of breaks (MAPS)

```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(grid)        # for unit()
# Optional but recommended for nice PNG/PDF outputs:
# install.packages(c("ragg","systemfonts"))
suppressWarnings({
  have_ragg <- requireNamespace("ragg", quietly = TRUE)
})

# ---- 0) INPUT expected: `dominant_breaks` ----
# Columns: country(ISO3), crop(BA/MA/RP/SB/SF/WT), break_year (numeric), ...
stopifnot(all(c("country","crop","break_year") %in% names(dominant_breaks)))

# ---- 1) Clean + label data ----
dom_clean <- dominant_breaks %>%
  filter(!is.na(crop), !is.na(break_year)) %>%
  mutate(
    crop_label = dplyr::recode(crop,
      "BA"="Barley","MA"="Maize","RP"="Rapeseed",
      "SB"="Soybean","SF"="Sunflower","WT"="Wheat",
      .default = crop
    ),
    decade = floor(break_year/10)*10,
    decade_lab = factor(paste0(decade,"s"))
  )

# Fix consistent decade order across facets/panels
dec_levels <- sort(unique(dom_clean$decade))
dom_clean <- dom_clean %>%
  mutate(decade_lab = factor(decade_lab, levels = paste0(dec_levels,"s")))

# ---- 2) World polygons + join ----
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%
  mutate(iso_fixed = dplyr::case_when(iso_a3 == "SDS" ~ "SSD", TRUE ~ iso_a3))

map_df <- world %>%
  left_join(dom_clean, by = c("iso_fixed" = "country"))

# Keep only rows with real crops for faceting (but we’ll draw ALL borders from `world`)
map_df_facets <- map_df %>% filter(!is.na(crop_label))

# ---- 3) Robinson projection (journal-friendly equal-area look)
world_robin <- st_transform(world, crs = "+proj=robin")
map_robin   <- st_transform(map_df_facets, crs = "+proj=robin")

# ---- 4) Helpers: consistent styling and two aesthetics (year vs decade) ----
year_limits <- range(map_robin$break_year, na.rm = TRUE)  # same scale across panels

base_map_bg <- function() {
  geom_sf(data = world_robin, fill = "grey96", color = "grey65", linewidth = 0.25)
}
base_theme <- theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    strip.text = element_text(face = "bold", size = 11),
    legend.position = "bottom",
    legend.key.width = unit(18, "pt"),
    legend.key.height = unit(6, "pt")
  )

# Continuous (year)
map_layer_year <- function(data_sub) {
  list(
    geom_sf(data = data_sub, aes(fill = break_year), color = "grey35", linewidth = 0.25),
    scale_fill_viridis_c(name = "Break year", limits = year_limits, na.value = "grey90")
  )
}
# Binned (decade)
map_layer_decade <- function(data_sub) {
  list(
    geom_sf(data = data_sub, aes(fill = decade_lab), color = "grey35", linewidth = 0.25),
    scale_fill_viridis_d(name = "Break decade", na.value = "grey90", drop = FALSE)
  )
}

# Split crops into two panels of three each
crops1 <- c("Barley","Maize","Rapeseed")
crops2 <- c("Soybean","Sunflower","Wheat")

# ---- 5) Build FOUR plots: (P1 year, P2 year) + (P1 decade, P2 decade) ----
p_year_1 <- ggplot() +
  base_map_bg() +
  map_layer_year(map_robin %>% filter(crop_label %in% crops1)) +
  coord_sf(crs = "+proj=robin", datum = NA) +
  facet_wrap(~ crop_label, ncol = 3) +
  labs(title = "Dominant structural break in yield–climate sensitivity — Panel 1 (by years)",
       subtitle = "Barley • Maize • Rapeseed") +
  base_theme

p_year_2 <- ggplot() +
  base_map_bg() +
  map_layer_year(map_robin %>% filter(crop_label %in% crops2)) +
  coord_sf(crs = "+proj=robin", datum = NA) +
  facet_wrap(~ crop_label, ncol = 3) +
  labs(title = "Dominant structural break in yield–climate sensitivity — Panel 2 (by years)",
       subtitle = "Soybean • Sunflower • Wheat") +
  base_theme

p_decade_1 <- ggplot() +
  base_map_bg() +
  map_layer_decade(map_robin %>% filter(crop_label %in% crops1)) +
  coord_sf(crs = "+proj=robin", datum = NA) +
  facet_wrap(~ crop_label, ncol = 3) +
  labs(title = "Dominant structural breaks by decade — Panel 1",
       subtitle = "Barley • Maize • Rapeseed") +
  base_theme

p_decade_2 <- ggplot() +
  base_map_bg() +
  map_layer_decade(map_robin %>% filter(crop_label %in% crops2)) +
  coord_sf(crs = "+proj=robin", datum = NA) +
  facet_wrap(~ crop_label, ncol = 3) +
  labs(title = "Dominant structural breaks by decade — Panel 2",
       subtitle = "Soybean • Sunflower • Wheat") +
  base_theme

# ---- 6) Show one (interactive dev) ----
# p_year_1; p_year_2; p_decade_1; p_decade_2

# ---- 7) Export (PDF vectors + high-res PNGs) ----
# Sizes ~7x4 in per panel tend to look good in journals; bump base_size if smaller
ggsave("dominant_breaks_year_panel1.pdf",   p_year_1,  width = 7, height = 4, device = cairo_pdf)
ggsave("dominant_breaks_year_panel2.pdf",   p_year_2,  width = 7, height = 4, device = cairo_pdf)
ggsave("dominant_breaks_decade_panel1.pdf", p_decade_1,width = 7, height = 4, device = cairo_pdf)
ggsave("dominant_breaks_decade_panel2.pdf", p_decade_2,width = 7, height = 4, device = cairo_pdf)

if (have_ragg) {
  ggsave("dominant_breaks_year_panel1.png",   p_year_1,  width = 7, height = 4, dpi = 450, device = ragg::agg_png)
  ggsave("dominant_breaks_year_panel2.png",   p_year_2,  width = 7, height = 4, dpi = 450, device = ragg::agg_png)
  ggsave("dominant_breaks_decade_panel1.png", p_decade_1,width = 7, height = 4, dpi = 450, device = ragg::agg_png)
  ggsave("dominant_breaks_decade_panel2.png", p_decade_2,width = 7, height = 4, dpi = 450, device = ragg::agg_png)
}
```

#4. Generate a collection of Summary Breaks statistics (for analytical purposes)

```{r}
# --- Structural Break Summary from segment_models_flat ---

# ----------------------------
# 0) Helpers & settings
# ----------------------------
tol <- 1e-6
sign_cat <- function(x, tol = 1e-6) {
  ifelse(x >  tol, "positive",
         ifelse(x < -tol, "negative", "no change"))
}

parse_me <- function(x) {
  # robust parser for "Tot_Precip=...; Total_GDD=...; yyy=..."
  if (is.na(x) || !nzchar(x))
    return(c(Total_GDD = NA_real_, Tot_Precip = NA_real_, yyy = NA_real_))
  parts <- str_split(x, ";")[[1]] %>% str_trim()
  kv <- str_split(parts, "=", n = 2, simplify = TRUE)
  if (ncol(kv) < 2)
    return(c(Total_GDD = NA_real_, Tot_Precip = NA_real_, yyy = NA_real_))
  vals <- suppressWarnings(as.numeric(kv[,2]))
  nm   <- kv[,1]
  out  <- setNames(rep(NA_real_, 3), c("Total_GDD","Tot_Precip","yyy"))
  idx  <- match(nm, names(out))
  out[idx[!is.na(idx)]] <- vals[!is.na(idx)]
  out
}

# ----------------------------
# 1) Parse marginal effects & normalize flag
# ----------------------------
coefs_long <- segment_models_flat %>%
  mutate(flag_before_after = if_else(flag_before_after %in% c("1", 1), 1, 0),
         me = map(marginal_effects, parse_me),
         Total_GDD  = map_dbl(me, ~ .x[["Total_GDD"]]),
         Tot_Precip = map_dbl(me, ~ .x[["Tot_Precip"]]),
         yyy        = map_dbl(me, ~ .x[["yyy"]])) %>%
  select(-me)

# ----------------------------
# 2) Coverage: detect complete vs. single-segment pairs
# ----------------------------
coverage_by_pair <- coefs_long %>%
  count(ppp, ccc, flag_before_after, name = "n_obs") %>%
  tidyr::pivot_wider(
    names_from = flag_before_after,
    values_from = n_obs,
    names_glue = "flag_{flag_before_after}",
    values_fill = 0
  ) %>%
  mutate(
    has_before = flag_0 > 0,
    has_after  = flag_1 > 0,
    pair_type = dplyr::case_when(
      has_before & has_after ~ "Complete (break detected)",
      has_before & !has_after ~ "Only 'before' segment",
      !has_before & has_after ~ "Only 'after' segment",
      TRUE ~ "Unknown"
    )
  )

# Per-crop coverage
coverage_summary <- coverage_by_pair %>%
  group_by(ppp, pair_type) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ppp) %>%
  mutate(share = round(100 * n / sum(n), 1)) %>%
  ungroup() %>%
  arrange(ppp, desc(n))

# Overall totals
coverage_total <- coverage_summary %>%
  group_by(pair_type) %>%
  summarise(
    total_n = sum(n),
    total_share = round(100 * total_n / sum(total_n), 1),
    .groups = "drop"
  )

# Country lists for mapping/inspection (no-breaks)
nobreak_countries <- coverage_by_pair %>%
  filter(pair_type != "Complete (break detected)") %>%
  mutate(which_segment = dplyr::case_when(
    pair_type == "Only 'before' segment" ~ "before_only",
    pair_type == "Only 'after' segment"  ~ "after_only",
    TRUE ~ "unknown"
  )) %>%
  select(ppp, ccc, which_segment) %>%
  arrange(ppp, ccc)

# ----------------------------
# 3) Direction & magnitudes on ALL pairs (includes "missing")
#    -> for completeness, if you want to show that ~X% are missing deltas pre-flood
# ----------------------------
coefs_wide_all <- coefs_long %>%
  tidyr::pivot_wider(
    id_cols = c(ccc, ppp),
    names_from = flag_before_after,
    values_from = c(Total_GDD, Tot_Precip, yyy),
    names_glue = "{.value}_{ifelse(flag_before_after==1,'after','before')}"
  )

deltas_all <- coefs_wide_all %>%
  mutate(
    d_GDD  = Total_GDD_after   - Total_GDD_before,
    d_Prec = Tot_Precip_after  - Tot_Precip_before,
    d_YYY  = yyy_after         - yyy_before
  )

dir_counts_all <- deltas_all %>%
  transmute(ppp,
            d_GDD_cat  = ifelse(is.finite(d_GDD),  sign_cat(d_GDD,  tol), "missing"),
            d_Prec_cat = ifelse(is.finite(d_Prec), sign_cat(d_Prec, tol), "missing"),
            d_YYY_cat  = ifelse(is.finite(d_YYY),  sign_cat(d_YYY,  tol), "missing")) %>%
  tidyr::pivot_longer(cols = starts_with("d_"), names_to = "variable", values_to = "direction") %>%
  mutate(variable = dplyr::recode(variable,
                                  d_GDD_cat  = "ΔGDD",
                                  d_Prec_cat = "ΔPrecip",
                                  d_YYY_cat  = "ΔYYY")) %>%
  group_by(ppp, variable, direction) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ppp, variable) %>%
  mutate(share = round(100 * n / sum(n), 1)) %>%
  ungroup()

# ----------------------------
# 4) Direction & magnitudes on COMPLETE pairs only (main stats)
# ----------------------------
coefs_wide_complete <- coverage_by_pair %>%
  filter(pair_type == "Complete (break detected)") %>%
  select(ccc, ppp) %>%
  inner_join(coefs_wide_all, by = c("ccc","ppp"))

deltas <- coefs_wide_complete %>%
  mutate(
    d_GDD  = Total_GDD_after   - Total_GDD_before,
    d_Prec = Tot_Precip_after  - Tot_Precip_before,
    d_YYY  = yyy_after         - yyy_before
  )

# Direction counts (no "missing" now)
dir_counts_complete <- deltas %>%
  transmute(ppp,
            d_GDD_cat  = sign_cat(d_GDD,  tol),
            d_Prec_cat = sign_cat(d_Prec, tol),
            d_YYY_cat  = sign_cat(d_YYY,  tol)) %>%
  tidyr::pivot_longer(cols = starts_with("d_"), names_to = "variable", values_to = "direction") %>%
  mutate(variable = dplyr::recode(variable,
                                  d_GDD_cat  = "ΔGDD",
                                  d_Prec_cat = "ΔPrecip",
                                  d_YYY_cat  = "ΔYYY")) %>%
  group_by(ppp, variable, direction) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ppp, variable) %>%
  mutate(share = round(100 * n / sum(n), 1)) %>%
  ungroup()

# Magnitudes & ratios
mag_summary_complete <- deltas %>%
  group_by(ppp) %>%
  summarise(
    mean_dGDD  = mean(d_GDD,  na.rm = TRUE),
    mean_dP    = mean(d_Prec, na.rm = TRUE),
    mean_dYYY  = mean(d_YYY,  na.rm = TRUE),
    abs_dGDD   = mean(abs(d_GDD),  na.rm = TRUE),
    abs_dP     = mean(abs(d_Prec), na.rm = TRUE),
    abs_dYYY   = mean(abs(d_YYY),  na.rm = TRUE),
    ratio_abs  = abs_dGDD / abs_dP,
    .groups = "drop"
  )

# Direction symbols by crop (mean sign across complete pairs)
dir_symbols_complete <- deltas %>%
  group_by(ppp) %>%
  summarise(
    GDD_dir  = case_when(mean(d_GDD,  na.rm=TRUE) >  tol ~ "↑ increase",
                         mean(d_GDD,  na.rm=TRUE) < -tol ~ "↓ decrease",
                         TRUE ~ "→ no change"),
    Prec_dir = case_when(mean(d_Prec, na.rm=TRUE) >  tol ~ "↑ increase",
                         mean(d_Prec, na.rm=TRUE) < -tol ~ "↓ decrease",
                         TRUE ~ "→ no change"),
    YYY_dir  = case_when(mean(d_YYY,  na.rm=TRUE) >  tol ~ "↑ increase",
                         mean(d_YYY,  na.rm=TRUE) < -tol ~ "↓ decrease",
                         TRUE ~ "→ no change"),
    .groups = "drop"
  )

# ----------------------------
# 5) Inspect in console (optional)
# ----------------------------
coverage_summary
coverage_total
nobreak_countries
dir_counts_all
dir_counts_complete
mag_summary_complete
dir_symbols_complete

# ----------------------------
# 6) Export
# ----------------------------
write.csv(coverage_summary,
          file.path(datpthout, paste0("BreakCoverage_ByCrop_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(coverage_total,
          file.path(datpthout, paste0("BreakCoverage_Total_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(nobreak_countries,
          file.path(datpthout, paste0("BreakCoverage_NoBreakCountries_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(dir_counts_all,
          file.path(datpthout, paste0("BreakDirection_Counts_ALL_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(dir_counts_complete,
          file.path(datpthout, paste0("BreakDirection_Counts_CompletePairs_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(mag_summary_complete,
          file.path(datpthout, paste0("BreakMagnitude_Summary_CompletePairs_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

write.csv(dir_symbols_complete,
          file.path(datpthout, paste0("BreakDirection_Symbols_CompletePairs_", format(Sys.Date(), "%Y%m%d"), ".csv")),
          row.names = FALSE)

cat("✅ Exported coverage, country lists, and break statistics to:", datpthout, "\n")
```

#5. Graphics of the breaks (Violin plots)

```{r}
# 1. Standardized crop color scheme (Nature-style: high contrast, colorblind friendly)
nice_cols <- c(
  "BA" = "#E69F00", "MA" = "#56B4E9", "RP" = "#F0E442",
  "SB" = "#009E73", "SF" = "#CC79A7", "WT" = "#0072B2"
)


# 1.a) Extract raw break years
bai_breaks <- purrr::imap_dfr(break_output$results, ~{
  parts <- strsplit(.y, "_")[[1]]
  tibble(
    country    = parts[1],
    crop       = parts[2],
    break_year = .x$break_years
  )
}) %>%
  unnest_longer(break_year) %>%
  filter(!is.na(break_year))   # drop NA rows here


# 2. Panel A: All Detected Candidates
p1 <- ggplot(bai_breaks, aes(x = crop, y = break_year, fill = crop)) +
  geom_violin(width = 0.9, alpha = 0.2, color = NA) +
  # Use quasirandom to avoid overplotting while showing every data point
  geom_quasirandom(aes(color = crop), size = 1.0, alpha = 0.4, width = 0.25) +
  # Black dots represent the median break year for that crop category
  stat_summary(fun = median, geom = "point", size = 2, color = "black") +
  scale_fill_manual(values = nice_cols) +
  scale_color_manual(values = nice_cols) +
  coord_flip() +
  labs(title = "a)   ", x = NULL, y = "Year") +
  theme_minimal(base_size = 11) + # Standard publication font size
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 12)
  )

# 3. Panel B: Dominant (Most Significant) Breaks
p2 <- ggplot(dominant_breaks, aes(x = crop, y = break_year, fill = crop)) +
  geom_violin(width = 0.9, alpha = 0.2, color = NA) +
  geom_quasirandom(aes(color = crop), size = 1.2, alpha = 0.6, width = 0.25) +
  # Black dots represent the median of the most significant break years
  stat_summary(fun = median, geom = "point", size = 2, color = "black") +
  scale_fill_manual(values = nice_cols) +
  scale_color_manual(values = nice_cols) +
  coord_flip() +
  labs(title = "b)    ", x = NULL, y = "Year") +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 12)
  )

# 4. Assemble with Patchwork
# Note: Nature figures often use specific widths (89mm or 183mm). 
# Adjust the width/height in ggsave accordingly.
combined_plot <- p1 / p2 + 
  plot_annotation(
    caption = "Black points indicate median values. Violins show density distribution of detected structural breaks.",
    theme = theme(plot.caption = element_text(size = 9, hjust = 0, italic = TRUE))
  )

print(combined_plot)

# 5. Export for Publication
ggsave("Figure_Breaks_Nature.pdf", combined_plot, width = 180, height = 220, units = "mm")

combined_plot <- p1 / p2 + 
  plot_annotation(
    caption = "Black points indicate median values. Violins show density distribution of detected structural breaks.",
    theme = theme(plot.caption = element_text(size = 9, hjust = 0, italic = TRUE))
  )

print(combined_plot)

# Export for publication
ggsave(
  filename = file.path(datpthout, "Figure_Breaks_Nature.png"),
  plot = combined_plot,
  width = 180,
  height = 220,
  units = "mm",
  dpi = 600,
  bg = "white"
)

```

## 6. Suplementary data output

```{r}
# =====================================================================
# SUPPLEMENTARY / ANNEX DATA EXPORT FOR PUBLICATION
# Generates:
#   Table S1 – Break detection summary
#   Table S2 – Regime comparison (levels, AMEs before/after)
#   Table S3 – Crop-level summary statistics
#   Table S4 – Heat-penalty direction counts
#   Annex 1 (PDF) – S1–S4 rendered from MEMORY (no CSV read-back)
# =====================================================================

library(dplyr)
library(readr)
library(grid)
library(gridExtra)

# Your deltas object is the regime-wide AME comparison table
test_df_3 <- deltas

out_dir <- file.path(datpthout, "Supplementary")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# ============================================================================
# TABLE S1 — Break detection summary (from break test)
# ============================================================================
table_S1_break_summary <- break_output$breaks_table %>%
  transmute(
    country = country,
    crop = crop,
    break_year = break_year,
    F_stat = F_stat,
    p_value = p_value,
    significant = significant
  ) %>%
  arrange(country, crop)

write_csv(table_S1_break_summary,
          file.path(out_dir, "Table_S1_Break_Summary.csv"))

# ============================================================================
# TABLE S2 — Regime comparison (LEVELS AMEs before vs after break)
# ============================================================================
table_S2_regime_levels <- test_df_3 %>%
  transmute(
    country = ccc,
    crop = ppp,
    AME_GDD_before  = Total_GDD_before,
    AME_GDD_after   = Total_GDD_after,
    d_GDD           = d_GDD,
    AME_Prec_before = Tot_Precip_before,
    AME_Prec_after  = Tot_Precip_after,
    d_Prec          = d_Prec,
    Trend_before    = yyy_before,
    Trend_after     = yyy_after,
    d_Trend         = d_YYY
  ) %>%
  arrange(country, crop)

write_csv(table_S2_regime_levels,
          file.path(out_dir, "Table_S2_Regime_Levels_AMEs.csv"))

# ============================================================================
# TABLE S3 — Crop-level summary (share stronger heat penalty)
# ============================================================================
table_S3_crop_summary <- test_df_3 %>%
  group_by(ppp) %>%
  summarise(
    n_pairs = sum(!is.na(d_GDD)),
    share_stronger_heat_penalty = mean(d_GDD < 0, na.rm = TRUE),
    share_weaker_heat_penalty   = mean(d_GDD > 0, na.rm = TRUE),
    median_d_GDD  = median(d_GDD, na.rm = TRUE),
    median_d_Prec = median(d_Prec, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(crop = ppp) %>%
  arrange(desc(share_stronger_heat_penalty))

write_csv(table_S3_crop_summary,
          file.path(out_dir, "Table_S3_Crop_Level_Summary.csv"))

# ============================================================================
# TABLE S4 — Direction counts (heat penalty classification)
# ============================================================================
table_S4_heat_direction <- test_df_3 %>%
  mutate(heat_response = case_when(
    is.na(d_GDD) ~ NA_character_,
    d_GDD < 0 ~ "Stronger heat penalty",
    d_GDD > 0 ~ "Weaker heat penalty",
    TRUE ~ "No change"
  )) %>%
  count(ppp, heat_response, name = "n") %>%
  rename(crop = ppp)

write_csv(table_S4_heat_direction,
          file.path(out_dir, "Table_S4_Heat_Penalty_Direction.csv"))

# ============================================================================
# ANNEX 1 (PDF): Supplementary break data tables — FROM MEMORY
# ============================================================================

stopifnot(
  exists("table_S1_break_summary"),
  exists("table_S2_regime_levels"),
  exists("table_S3_crop_summary"),
  exists("table_S4_heat_direction")
)

# ---- helper: render a (possibly long) df across multiple PDF pages ----
df_to_pages <- function(df, title, note = NULL,
                        max_rows = 40, digits = 4, font = 7) {

  df_print <- df
  num_cols <- vapply(df_print, is.numeric, logical(1))
  df_print[num_cols] <- lapply(df_print[num_cols], function(x) round(x, digits))

  n <- nrow(df_print)
  if (n == 0) {
    grid.newpage()
    grid.text(title, x = 0.02, y = 0.98, just = c("left","top"),
              gp = gpar(fontface = "bold", fontsize = 12))
    grid.text("(No rows)", x = 0.02, y = 0.92, just = c("left","top"),
              gp = gpar(fontsize = 10, col = "grey35"))
    return(invisible(NULL))
  }

  starts <- seq(1, n, by = max_rows)

  for (k in seq_along(starts)) {
    i1 <- starts[k]
    i2 <- min(i1 + max_rows - 1, n)
    chunk <- df_print[i1:i2, , drop = FALSE]

    tg <- tableGrob(
      chunk, rows = NULL,
      theme = ttheme_minimal(
        core   = list(fg_params = list(cex = font/10)),
        colhead= list(fg_params = list(cex = (font+1)/10, fontface = "bold"))
      )
    )

    grid.newpage()
    grid.text(
      if (length(starts) > 1) paste0(title, " (rows ", i1, "–", i2, " of ", n, ")") else title,
      x = 0.02, y = 0.98, just = c("left","top"),
      gp = gpar(fontface = "bold", fontsize = 12)
    )

    if (!is.null(note)) {
      grid.text(note, x = 0.02, y = 0.945, just = c("left","top"),
                gp = gpar(fontsize = 9, col = "grey25"))
    }

    # generous viewport; table will paginate rather than clip horizontally
    pushViewport(viewport(x = 0.5, y = 0.46, width = 0.98, height = 0.86))
    grid.draw(tg)
    popViewport()
  }

  invisible(NULL)
}

# ---- output file ----
annex_file <- file.path(out_dir, "Annex_1_Supplementary_Break_Tables.pdf")
pdf(annex_file, width = 11, height = 8.5)  # landscape

# ---- cover page ----
grid.newpage()
grid.text("Annex 1 — Supplementary Break Data Tables",
          x = 0.5, y = 0.62, gp = gpar(fontface = "bold", fontsize = 18))
grid.text(paste("Generated:", Sys.Date()),
          x = 0.5, y = 0.54, gp = gpar(fontsize = 11))
grid.text("Tables S1–S4 generated directly from the replication pipeline.",
          x = 0.5, y = 0.42, gp = gpar(fontsize = 10))

# ---- S1 ----
df_to_pages(
  table_S1_break_summary,
  "Table S1 — Break detection summary",
  "Candidate breaks with F-statistic, p-value, and significance flag.",
  max_rows = 45, font = 7
)

# ---- S2 (SPLIT WIDE TABLE INTO TWO PAGES BY COLUMNS) ----
S2a <- table_S2_regime_levels %>%
  dplyr::select(country, crop, AME_GDD_before, AME_GDD_after, d_GDD, Trend_before, Trend_after, d_Trend)

S2b <- table_S2_regime_levels %>%
  dplyr::select(country, crop, AME_Prec_before, AME_Prec_after, d_Prec, Trend_before, Trend_after, d_Trend)

df_to_pages(
  S2a,
  "Table S2a — Regime comparison (levels): GDD",
  "Average marginal effects before/after break; levels-based comparison.",
  max_rows = 45, font = 7
)

df_to_pages(
  S2b,
  "Table S2b — Regime comparison (levels): Precipitation",
  "Average marginal effects before/after break; levels-based comparison.",
  max_rows = 45, font = 7
)

# ---- S3 ----
df_to_pages(
  table_S3_crop_summary,
  "Table S3 — Crop-level heat penalty summary",
  "Share of series with stronger post-break heat penalty; median ΔGDD and ΔPrecip.",
  max_rows = 60, font = 7
)

# ---- S4 ----
df_to_pages(
  table_S4_heat_direction,
  "Table S4 — Heat penalty direction counts",
  "Counts of stronger / weaker / unchanged post-break heat penalty by crop.",
  max_rows = 60, font = 7
)

dev.off()

cat("✅ Supplementary tables exported to:", out_dir, "\n")
cat("✅ Annex 1 PDF generated:", annex_file, "\n")
```

### --- EXTRA SANITY CHECKS ---

```{r}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(broom)

pstars <- function(p) dplyr::case_when(
  is.na(p) ~ "",
  p < 0.01 ~ "***",
  p < 0.05 ~ "**",
  p < 0.1  ~ "*",
  TRUE ~ ""
)

get_term <- function(tidy_df, term_name) {
  out <- tidy_df %>% filter(term == term_name)
  if (nrow(out) == 0) return(tibble(estimate = NA_real_, std.error = NA_real_, p.value = NA_real_))
  out %>% select(estimate, std.error, p.value)
}


target_country <- "FRA"
france_crops   <- c("BA","MA","RP","WT","SB")

# dominant break per crop (Chow-significant highest F)
fr_breaks <- break_output$breaks_table %>%
  filter(country == target_country, crop %in% france_crops, significant) %>%
  group_by(crop) %>%
  slice_max(F_stat, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(crop, break_year)

# keep all crops; those w/o break will be NA
fr_breaks <- tibble(crop = france_crops) %>%
  left_join(fr_breaks, by = "crop")

fr_ts <- yield_climate_data_cleaned %>%
  filter(ccc == target_country, ppp %in% france_crops) %>%
  mutate(ppp = as.character(ppp))

fit_crop_interaction <- function(crop_code, byear) {
  
  df <- fr_ts %>% filter(ppp == crop_code) %>% arrange(yyy)
  
  if (is.na(byear)) {
    # no break: fit baseline model only (no Δ)
    mod <- lm(YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy, data = df)
    td  <- broom::tidy(mod)
    
    gdd_b  <- get_term(td, "Total_GDD")
    prc_b  <- get_term(td, "Tot_Precip")
    
    return(tibble(
      Crop = crop_code, Break = NA_integer_,
      me_gdd_before = gdd_b$estimate,
      me_gdd_after  = NA_real_,
      d_me_gdd      = NA_real_, p_d_me_gdd = NA_real_, lo_d_me_gdd = NA_real_, hi_d_me_gdd = NA_real_,
      me_prec_before = prc_b$estimate,
      me_prec_after  = NA_real_,
      d_me_prec      = NA_real_, p_d_me_prec = NA_real_, lo_d_me_prec = NA_real_, hi_d_me_prec = NA_real_
    ))
  }
  
  df <- df %>% mutate(post = as.integer(yyy >= byear))
  
  mod <- lm(
    YLD ~ (Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy) * post,
    data = df
  )
  
  td <- broom::tidy(mod)
  
  # BEFORE slopes
  gdd_b <- get_term(td, "Total_GDD")
  prc_b <- get_term(td, "Tot_Precip")
  
  # DELTAS are interaction slopes (After - Before)
  gdd_d <- get_term(td, "Total_GDD:post")
  prc_d <- get_term(td, "Tot_Precip:post")
  
  # 95% CI for deltas
  gdd_lo <- gdd_d$estimate - 1.96 * gdd_d$std.error
  gdd_hi <- gdd_d$estimate + 1.96 * gdd_d$std.error
  
  prc_lo <- prc_d$estimate - 1.96 * prc_d$std.error
  prc_hi <- prc_d$estimate + 1.96 * prc_d$std.error
  
  tibble(
    Crop = crop_code, Break = as.integer(byear),
    
    me_gdd_before = gdd_b$estimate,
    me_gdd_after  = gdd_b$estimate + gdd_d$estimate,
    d_me_gdd      = gdd_d$estimate,
    p_d_me_gdd    = gdd_d$p.value,
    lo_d_me_gdd   = gdd_lo,
    hi_d_me_gdd   = gdd_hi,
    
    me_prec_before = prc_b$estimate,
    me_prec_after  = prc_b$estimate + prc_d$estimate,
    d_me_prec      = prc_d$estimate,
    p_d_me_prec    = prc_d$p.value,
    lo_d_me_prec   = prc_lo,
    hi_d_me_prec   = prc_hi
  )
}

fr_table_raw <- fr_breaks %>%
  mutate(tab = map2(crop, break_year, fit_crop_interaction)) %>%
  select(tab) %>%
  unnest(tab)



fr_table_show <- fr_table_raw %>%
  transmute(
    Crop,
    Break = ifelse(is.na(Break), "—", as.character(Break)),
    
    `ME(GDD) Before`  = signif(me_gdd_before, 3),
    `ME(GDD) After`   = signif(me_gdd_after, 3),
    `ΔME(GDD)`        = ifelse(is.na(d_me_gdd), "NA",
                               paste0(signif(d_me_gdd, 3), pstars(p_d_me_gdd))),
    
    `ME(Prec) Before` = signif(me_prec_before, 3),
    `ME(Prec) After`  = signif(me_prec_after, 3),
    `ΔME(Prec)`       = ifelse(is.na(d_me_prec), "NA",
                               paste0(signif(d_me_prec, 3), pstars(p_d_me_prec)))
  )

fr_table_show


library(flextable)

ft <- flextable(fr_table_show) %>%
  autofit() %>%
  fontsize(size = 9, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "all")

ft



delta_long <- fr_table_raw %>%
  select(Crop, Break,
         d_me_gdd, lo_d_me_gdd, hi_d_me_gdd, p_d_me_gdd,
         d_me_prec, lo_d_me_prec, hi_d_me_prec, p_d_me_prec) %>%
  pivot_longer(
    cols = -c(Crop, Break),
    names_to = c(".value","var"),
    names_pattern = "(d_me|lo_d_me|hi_d_me|p_d_me)_(gdd|prec)"
  ) %>%
  mutate(
    variable = recode(var, gdd = "ΔME(GDD)", prec = "ΔME(Precip)"),
    sig_cat = case_when(
      p_d_me < 0.05 ~ "p<0.05",
      p_d_me < 0.1  ~ "p<0.10",
      TRUE ~ "n.s."
    ),
    Crop = factor(Crop, levels = france_crops)
  )

p_delta <- ggplot(delta_long, aes(x = d_me, y = Crop)) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  geom_pointrange(aes(xmin = lo_d_me, xmax = hi_d_me, shape = sig_cat),
                  linewidth = 0.5) +
  facet_wrap(~variable, scales = "free_x") +
  labs(
    title = "France: regime shift in climate sensitivity at dominant breakpoints",
    subtitle = "Δ marginal effects (After − Before); 95% CI from interaction term",
    x = "Δ marginal effect (After − Before)",
    y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        legend.position = "bottom")

print(p_delta)

ggsave(file.path(datpthout, "Fig_France_DeltaME_Dotplots.pdf"),
       p_delta, width = 180, height = 95, units = "mm", device = cairo_pdf)
ggsave(file.path(datpthout, "Fig_France_DeltaME_Dotplots.png"),
       p_delta, width = 180, height = 95, units = "mm", dpi = 600)





```

#sanity With and without SUPf

```{r}
# ============================================================
# SUPF vs NO-SUPF COMPARISON WORKFLOW (FULL DIAGNOSTIC BLOCK)
# ============================================================

library(dplyr)

# -------------------------------
# 1. Run both specifications
# -------------------------------
out_nosupf <- run_break_analysis(
  data = yield_climate_data_cleaned,
  selected_crops = selected_crops,
  selected_countries = selected_countries,
  model_formula = "YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy",
  window_years = 15,
  global_test = FALSE
)

out_supf <- run_break_analysis(
  data = yield_climate_data_cleaned,
  selected_crops = selected_crops,
  selected_countries = selected_countries,
  model_formula = "YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy",
  window_years = 15,
  global_test = TRUE
)

# -------------------------------
# 2. Panel coverage comparison
# -------------------------------
nosupf_pairs <- out_nosupf$breaks_summary %>% select(country, crop)
supf_pairs   <- out_supf$breaks_summary   %>% select(country, crop)

filtered_out <- anti_join(nosupf_pairs, supf_pairs, by = c("country","crop"))
common_pairs <- inner_join(nosupf_pairs, supf_pairs, by = c("country","crop"))

coverage_summary <- tibble(
  total_panels_nosupf = nrow(nosupf_pairs),
  total_panels_supf   = nrow(supf_pairs),
  panels_removed_by_supf = nrow(filtered_out),
  share_removed_pct = round(100 * nrow(filtered_out) / nrow(nosupf_pairs), 1)
)

# -------------------------------
# 3. Compare dominant break years
# -------------------------------
cmp <- inner_join(
  out_nosupf$breaks_summary %>%
    select(country, crop,
           break_year_nosupf = year_max_F,
           chow_p_nosupf = chow_p_value,
           global_p_nosupf = global_p_value,
           supF_year_nosupf = supF_year),
  out_supf$breaks_summary %>%
    select(country, crop,
           break_year_supf = year_max_F,
           chow_p_supf = chow_p_value,
           global_p_supf = global_p_value,
           supF_year_supf = supF_year),
  by = c("country","crop")
)

break_year_change_summary <- tibble(
  panels_compared = nrow(cmp),
  break_year_changed = sum(cmp$break_year_nosupf != cmp$break_year_supf, na.rm = TRUE)
)

# -------------------------------
# 4. Classify break strength (Chow)
# -------------------------------
classify_strength <- function(p) {
  case_when(
    is.na(p) ~ "NA",
    p <= 0.01 ~ "Strong",
    p <= 0.05 ~ "Moderate",
    p <= 0.10 ~ "Weak",
    TRUE ~ "Not significant"
  )
}

break_strength_table <- out_nosupf$breaks_summary %>%
  mutate(
    chow_strength = classify_strength(chow_p_value),
    chow_sig_05 = !is.na(chow_p_value) & chow_p_value <= 0.05
  ) %>%
  count(chow_strength, name = "n_panels")

# -------------------------------
# 5. Global vs Local support flags
# -------------------------------
support_flags_table <- out_nosupf$breaks_summary %>%
  mutate(
    global_supported = !is.na(global_p_value) & global_p_value <= 0.05,
    local_supported  = !is.na(chow_p_value) & chow_p_value <= 0.05,
    support_class = case_when(
      global_supported & local_supported ~ "Global + Local",
      global_supported & !local_supported ~ "Global only",
      !global_supported & local_supported ~ "Local only",
      TRUE ~ "Neither"
    )
  ) %>%
  count(support_class, name = "n_panels")

# -------------------------------
# 6. Diagnostic disagreement table
# -------------------------------
diagnostic_table <- out_nosupf$breaks_summary %>%
  mutate(
    global_supported = !is.na(global_p_value) & global_p_value <= 0.05,
    local_supported  = !is.na(chow_p_value) & chow_p_value <= 0.05,
    support_class = case_when(
      global_supported & local_supported ~ "Global + Local",
      global_supported & !local_supported ~ "Global only",
      !global_supported & local_supported ~ "Local only",
      TRUE ~ "Neither"
    )
  ) %>%
  select(
    country, crop,
    supF_year, supF_stat, global_p_value,
    year_max_F, max_F_stat, chow_p_value,
    support_class
  ) %>%
  arrange(desc(support_class), chow_p_value)

disagreement_only <- diagnostic_table %>%
  filter(support_class %in% c("Global only", "Local only"))

# -------------------------------
# 7. Print compact summary
# -------------------------------
cat("\n==============================\n")
cat("SUPF vs NO-SUPF COMPARISON\n")
cat("==============================\n")

cat("\nPanel Coverage:\n")
print(coverage_summary)

cat("\nBreak-Year Differences:\n")
print(break_year_change_summary)

cat("\nBreak Strength Distribution (Chow):\n")
print(break_strength_table)

cat("\nGlobal vs Local Support:\n")
print(support_flags_table)

cat("\nPanels Removed by supF (first 20):\n")
print(head(filtered_out, 20))

cat("\nDisagreement Table (first 20 rows):\n")
print(head(disagreement_only, 20))

# Optional: return all objects for further use
comparison_results <- list(
  coverage_summary = coverage_summary,
  break_year_change_summary = break_year_change_summary,
  break_strength_table = break_strength_table,
  support_flags_table = support_flags_table,
  filtered_out_panels = filtered_out,
  diagnostic_table = diagnostic_table,
  disagreement_only = disagreement_only
)

```

# Expanding on the sanity

```{r}

# ============================================================
# BREAK QUALITY & ROBUSTNESS ANALYSIS
# ============================================================

library(dplyr)
library(ggplot2)
library(scales)

df <- out_nosupf$breaks_summary

# ------------------------------------------------------------
# 1. Add strength + support classification
# ------------------------------------------------------------
df2 <- df %>%
  mutate(
    chow_strength = case_when(
      is.na(chow_p_value) ~ "NA",
      chow_p_value <= 0.01 ~ "Strong",
      chow_p_value <= 0.05 ~ "Moderate",
      chow_p_value <= 0.10 ~ "Weak",
      TRUE ~ "Not significant"
    ),
    global_supported = !is.na(global_p_value) & global_p_value <= 0.05,
    local_supported  = !is.na(chow_p_value) & chow_p_value <= 0.05,
    support_class = case_when(
      global_supported & local_supported ~ "Global + Local",
      global_supported & !local_supported ~ "Global only",
      !global_supported & local_supported ~ "Local only",
      TRUE ~ "Neither"
    )
  )

# ------------------------------------------------------------
# 2. TABLE: Global+Local vs others by crop
# ------------------------------------------------------------
table_support_crop <- df2 %>%
  mutate(class2 = if_else(support_class == "Global + Local", "Global+Local", "Other")) %>%
  count(crop, class2) %>%
  tidyr::pivot_wider(names_from = class2, values_from = n, values_fill = 0) %>%
  mutate(
    total = `Global+Local` + Other,
    share_global_local = round(100 * `Global+Local` / total, 1)
  ) %>%
  arrange(desc(share_global_local))

table_support_crop
# ---- This is your paper-ready table ----

# ------------------------------------------------------------
# 3. FIGURE: Share of Strong vs Weak breaks by crop
# ------------------------------------------------------------
fig_strength <- df2 %>%
  filter(chow_strength %in% c("Strong","Moderate","Weak")) %>%
  mutate(group = if_else(chow_strength == "Strong", "Strong", "Weak/Moderate")) %>%
  count(crop, group) %>%
  group_by(crop) %>%
  mutate(share = n / sum(n)) %>%
  ungroup()

p_strength <- ggplot(fig_strength, aes(x = crop, y = share, fill = group)) +
  geom_col(position = "stack", width = 0.7) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("Strong" = "#1b9e77", "Weak/Moderate" = "#d95f02")) +
  labs(
    title = "Strength of detected structural breaks by crop",
    x = "Crop",
    y = "Share of breaks",
    fill = "Break strength"
  ) +
  theme_bw(base_size = 11)

print(p_strength)

# ------------------------------------------------------------
# 4. BREAK CREDIBILITY INDEX
# ------------------------------------------------------------
# Combines:
#   - Chow F-stat magnitude
#   - Chow p-value
#   - Global support
#
# Normalization makes index comparable across panels.

df_index <- df2 %>%
  mutate(
    F_norm = max_F_stat / max(max_F_stat, na.rm = TRUE),
    p_norm = 1 - pmin(chow_p_value, 0.10) / 0.10,
    support_bonus = case_when(
      support_class == "Global + Local" ~ 1,
      support_class == "Local only" ~ 0.7,
      support_class == "Global only" ~ 0.5,
      TRUE ~ 0.2
    ),
    credibility_index = round(100 * (0.5 * F_norm + 0.3 * p_norm + 0.2 * support_bonus), 1)
  )

# Summary distribution
credibility_summary <- df_index %>%
  summarise(
    mean_index = mean(credibility_index, na.rm = TRUE),
    median_index = median(credibility_index, na.rm = TRUE),
    min_index = min(credibility_index, na.rm = TRUE),
    max_index = max(credibility_index, na.rm = TRUE)
  )

credibility_summary

# Top and bottom examples
top_breaks <- df_index %>%
  arrange(desc(credibility_index)) %>%
  select(country, crop, year_max_F, credibility_index) %>%
  head(10)

low_breaks <- df_index %>%
  arrange(credibility_index) %>%
  select(country, crop, year_max_F, credibility_index) %>%
  head(10)

top_breaks
low_breaks

# Optional: histogram
ggplot(df_index, aes(x = credibility_index)) +
  geom_histogram(bins = 25, fill = "#2c7fb8", color = "white") +
  labs(
    title = "Distribution of Break Credibility Index",
    x = "Credibility index (0–100)",
    y = "Number of panels"
  ) +
  theme_bw(base_size = 11)

```


```{r}
# ============================================================
# FRANCE MAIZE (MA) LEVELS DIAGNOSTIC PLOT
# - observed yields
# - fitted values before and after the dominant break
# - break year and segment R² shown on the figure
# ============================================================

library(dplyr)
library(ggplot2)

# ----------------------------
# 1. Settings
# ----------------------------
target_country <- "FRA"
target_crop    <- "MA"

base_formula <- YLD ~ Total_GDD + I(Total_GDD^2) + Tot_Precip + I(Tot_Precip^2) + yyy

# ----------------------------
# 2. Get dominant break year
# ----------------------------
fr_ma_break <- break_output$breaks_table %>%
  filter(country == target_country, crop == target_crop, significant) %>%
  slice_max(F_stat, n = 1, with_ties = FALSE)

stopifnot(nrow(fr_ma_break) == 1)

break_year <- fr_ma_break$break_year[[1]]

# ----------------------------
# 3. Prepare data
# ----------------------------
df_ma <- yield_climate_data_cleaned %>%
  filter(ccc == target_country, ppp == target_crop) %>%
  arrange(yyy) %>%
  mutate(regime = if_else(yyy < break_year, "Before", "After"))

stopifnot(nrow(df_ma) > 0)

# ----------------------------
# 4. Fit separate levels models
# ----------------------------
df_before <- df_ma %>% filter(yyy < break_year)
df_after  <- df_ma %>% filter(yyy >= break_year)

mod_before <- lm(base_formula, data = df_before)
mod_after  <- lm(base_formula, data = df_after)

df_before <- df_before %>%
  mutate(fitted = predict(mod_before, newdata = df_before))

df_after <- df_after %>%
  mutate(fitted = predict(mod_after, newdata = df_after))

plot_df <- bind_rows(df_before, df_after)

r2_before <- summary(mod_before)$r.squared
r2_after  <- summary(mod_after)$r.squared

# ----------------------------
# 5. Label positions
# ----------------------------
y_top <- max(plot_df$YLD, plot_df$fitted, na.rm = TRUE)
y_bot <- min(plot_df$YLD, plot_df$fitted, na.rm = TRUE)
y_rng <- y_top - y_bot

ann_df <- tibble(
  x = c(min(df_before$yyy, na.rm = TRUE) + 2,
        max(df_after$yyy, na.rm = TRUE) - 2),
  y = c(y_top - 0.08 * y_rng,
        y_top - 0.50 * y_rng),
  lab = c(
    paste0("Before break\nR² = ", sprintf("%.3f", r2_before)),
    paste0("After break\nR² = ", sprintf("%.3f", r2_after))
  ),
  hjust = c(0, 1)
)

# ----------------------------
# 6. Plot
# ----------------------------
p_fr_ma <- ggplot(plot_df, aes(x = yyy)) +
  # observed series
  geom_line(aes(y = YLD), linewidth = 0.7, color = "grey35") +
  geom_point(aes(y = YLD), size = 1.6, color = "grey20") +

  # fitted values by regime
  geom_line(
    data = df_before,
    aes(y = fitted),
    linewidth = 1.2,
    color = "#1f78b4"
  ) +
  geom_line(
    data = df_after,
    aes(y = fitted),
    linewidth = 1.2,
    color = "#d95f02"
  ) +

  # break line
  geom_vline(
    xintercept = break_year,
    linetype = "dashed",
    linewidth = 0.7,
    color = "black"
  ) +

  # break year label
  annotate(
    "text",
    x = break_year + 0.5,
    y = y_top - 0.02 * y_rng,
    label = as.character(break_year),
    hjust = 0,
    vjust = 1,
    size = 4
  ) +

  # R2 annotations
  geom_label(
    data = ann_df,
    aes(x = x, y = y, label = lab, hjust = hjust),
    inherit.aes = FALSE,
    size = 3.4,
    label.size = 0.2,
    fill = "white"
  ) +

  scale_x_continuous(
    breaks = pretty(plot_df$yyy, n = 6),
    expand = expansion(mult = c(0.02, 0.06))
  ) +
  labs(
    title = "France maize: levels-based structural break diagnostic",
    subtitle = paste0(
      "Observed yields with regime-specific fitted values; dominant break year = ",
      break_year
    ),
    x = "Year",
    y = "Yield (t/ha)",
    caption = "Grey = observed yields; blue = pre-break fitted values; orange = post-break fitted values."
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10),
    panel.grid.minor = element_blank(),
    plot.caption = element_text(size = 9, hjust = 0),
    axis.title = element_text(face = "bold")
  )

print(p_fr_ma)

# ----------------------------
# 7. Save high-res PNG
# ----------------------------
ggsave(
  filename = file.path(datpthout, "France_Maize_Levels_Break_Diagnostic.png"),
  plot = p_fr_ma,
  width = 180,
  height = 110,
  units = "mm",
  dpi = 600,
  bg = "white"
)
```

```{r}
plot_france_levels_break <- function(crop_code) {
  br <- break_output$breaks_table %>%
    filter(country == "FRA", crop == crop_code, significant) %>%
    slice_max(F_stat, n = 1, with_ties = FALSE)

  if (nrow(br) == 0) return(NULL)

  break_year <- br$break_year[[1]]

  df_crop <- yield_climate_data_cleaned %>%
    filter(ccc == "FRA", ppp == crop_code) %>%
    arrange(yyy)

  df_before <- df_crop %>% filter(yyy < break_year)
  df_after  <- df_crop %>% filter(yyy >= break_year)

  if (nrow(df_before) < 10 || nrow(df_after) < 10) return(NULL)

  mod_before <- lm(base_formula, data = df_before)
  mod_after  <- lm(base_formula, data = df_after)

  df_before <- df_before %>% mutate(fitted = predict(mod_before, newdata = df_before))
  df_after  <- df_after  %>% mutate(fitted = predict(mod_after,  newdata = df_after))

  plot_df <- bind_rows(df_before, df_after)

  r2_before <- summary(mod_before)$r.squared
  r2_after  <- summary(mod_after)$r.squared

  y_top <- max(plot_df$YLD, plot_df$fitted, na.rm = TRUE)
  y_bot <- min(plot_df$YLD, plot_df$fitted, na.rm = TRUE)
  y_rng <- y_top - y_bot

  ann_df <- tibble(
    x = c(min(df_before$yyy, na.rm = TRUE) + 2,
          max(df_after$yyy, na.rm = TRUE) - 2),
    y = c(y_top - 0.08 * y_rng,
          y_top - 0.50 * y_rng),
    lab = c(
      paste0("Before break\nR² = ", sprintf("%.3f", r2_before)),
      paste0("After break\nR² = ", sprintf("%.3f", r2_after))
    ),
    hjust = c(0, 1)
  )

  ggplot(plot_df, aes(x = yyy)) +
    geom_line(aes(y = YLD), linewidth = 0.7, color = "grey35") +
    geom_point(aes(y = YLD), size = 1.6, color = "grey20") +
    geom_line(data = df_before, aes(y = fitted), linewidth = 1.2, color = "#1f78b4") +
    geom_line(data = df_after,  aes(y = fitted), linewidth = 1.2, color = "#d95f02") +
    geom_vline(xintercept = break_year, linetype = "dashed", linewidth = 0.7, color = "black") +
    annotate(
      "text",
      x = break_year + 0.5,
      y = y_top - 0.02 * y_rng,
      label = as.character(break_year),
      hjust = 0, vjust = 1, size = 4
    ) +
    geom_label(
      data = ann_df,
      aes(x = x, y = y, label = lab, hjust = hjust),
      inherit.aes = FALSE,
      size = 3.4,
      linewidth = 0.2,
      fill = "white"
    ) +
    scale_x_continuous(
      breaks = pretty(plot_df$yyy, n = 6),
      expand = expansion(mult = c(0.02, 0.06))
    ) +
    labs(
      title = paste0("France ", crop_code, ": levels-based structural break diagnostic"),
      subtitle = paste0("Observed yields with regime-specific fitted values; dominant break year = ", break_year),
      x = "Year",
      y = "Yield (t/ha)"
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face = "bold")
    )
}
france_break_plot_crops <- fr_breaks %>%
  filter(has_break) %>%
  pull(crop)

france_break_plots <- purrr::set_names(
  purrr::map(france_break_plot_crops, plot_france_levels_break),
  france_break_plot_crops
)

purrr::iwalk(france_break_plots, ~ {
  if (!is.null(.x)) {
    ggsave(
      filename = file.path(datpthout, paste0("France_", .y, "_Levels_Break_Diagnostic.png")),
      plot = .x,
      width = 180, height = 110, units = "mm", dpi = 600, bg = "white"
    )
  }
})
```

