
---
title: "Climate sensitivity — FD panel & robustness (clean)"
output: html_document
date: "2026-03-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r libraries}
# Core analysis stack used in the paper:
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(readr)

library(plm)
library(broom)
library(marginaleffects)

library(ggplot2)

# Only needed if you export Annex PDF
# library(grid)
# library(gridExtra)
```

```{r paths}
datpthout <- "C:\\Users\\arsovsp\\Documents\\Climate Risk\\output\\"
```

## 1) Load the latest processed breaks dataset

```{r load-data}
files <- list.files(
  path = datpthout,
  pattern = "Processed_data_breaks_flagged",
  full.names = TRUE
)

stopifnot(length(files) > 0)
info <- file.info(files)
latest_file <- files[which.max(info$mtime)]

sensitivity_dataset <- readr::read_csv(latest_file, show_col_types = FALSE)

# Paper convention: break_year == 2022 (or NA) treated as "NO_BREAK"
sensitivity_tagged <- sensitivity_dataset %>%
  mutate(
    break_tag = if_else(break_year == 2022 | is.na(break_year), "NO_BREAK", "HAS_BREAK")
  )

# Baseline FD sample: drop synthetic/no-break panels
panel_dataset_fd <- sensitivity_tagged %>%
  filter(break_tag == "HAS_BREAK") %>%
  mutate(
    flag_before_after = factor(as.integer(flag_before_after), levels = c(0, 1)),
    ccc = as.character(ccc),
    ppp = as.character(ppp)
  )

# Stress tests: deliberately misclassify NO_BREAK panels into a regime
panel_dataset_flood_after <- sensitivity_tagged %>%
  mutate(
    flag_before_after = if_else(break_tag == "NO_BREAK", 1L, as.integer(flag_before_after)),
    flag_before_after = factor(flag_before_after, levels = c(0, 1)),
    ccc = as.character(ccc),
    ppp = as.character(ppp)
  )

panel_dataset_flood_before <- sensitivity_tagged %>%
  mutate(
    flag_before_after = if_else(break_tag == "NO_BREAK", 0L, as.integer(flag_before_after)),
    flag_before_after = factor(flag_before_after, levels = c(0, 1)),
    ccc = as.character(ccc),
    ppp = as.character(ppp)
  )

# Placebo: remove regime split entirely (all "before")
panel_dataset_placebo <- sensitivity_tagged %>%
  mutate(
    flag_before_after = factor(0L, levels = c(0, 1)),
    ccc = as.character(ccc),
    ppp = as.character(ppp)
  )

date_str <- format(Sys.Date(), "%Y%m%d")
crops_fd <- sort(unique(panel_dataset_fd$ppp))
```

## 2) Estimation helpers (paper specification)

```{r helpers}
# Cluster-robust SEs by country (group) — as per Methods
vcov_cluster_ccc <- function(model) {
  plm::vcovHC(model, type = "HC1", cluster = "group")
}

# FD-Pool with regime interaction (baseline specification)
fit_fd_pool <- function(df, crop_id) {
  d <- df %>% filter(ppp == crop_id)
  if (nrow(d) < 10) return(NULL)
  if (all(d$flag_before_after == "0") || all(d$flag_before_after == "1")) return(NULL)

  mod <- plm::plm(
    Delta_Yield ~ (Delta_GDD + I(Delta_GDD^2) + Delta_Precip + I(Delta_Precip^2)) * flag_before_after,
    data  = d,
    index = c("ccc", "yyy"),
    model = "within"
  )

  list(model = mod, data = d)
}

# AMEs by regime over the full crop sample (avg_slopes with by=)
extract_ames_by_regime <- function(fit_obj, crop_id, model_label) {
  if (is.null(fit_obj)) return(NULL)

  am <- marginaleffects::avg_slopes(
    fit_obj$model,
    variables = c("Delta_GDD", "Delta_Precip"),
    by        = "flag_before_after",
    newdata   = fit_obj$data,
    vcov      = vcov_cluster_ccc
  ) %>%
    as.data.frame() %>%
    mutate(
      crop = crop_id,
      model_type = model_label,
      period = if_else(as.character(flag_before_after) == "0", "Before Break", "Post Break")
    ) %>%
    select(model_type, crop, period, term, estimate, std.error, conf.low, conf.high, p.value)

  am
}

# Coefficients + within R2 (for tables / annex)
extract_coefs <- function(fit_obj, crop_id, model_label) {
  if (is.null(fit_obj)) return(NULL)

  sm <- summary(fit_obj$model)
  r2_within <- unname(sm$r.squared["within"])

  broom::tidy(fit_obj$model, vcov = vcov_cluster_ccc) %>%
    mutate(
      crop = crop_id,
      model_type = model_label,
      r2_within = r2_within
    )
}

# Rescale AMEs to "per 100 ΔGDD / per 100 mm ΔPrecip" (Figure 4 convention)
rescale_ames_100 <- function(df) {
  df %>%
    mutate(
      scale = case_when(
        term == "Delta_GDD"    ~ 100,
        term == "Delta_Precip" ~ 100,
        TRUE ~ 1
      ),
      estimate  = estimate  * scale,
      std.error = std.error * scale,
      conf.low  = conf.low  * scale,
      conf.high = conf.high * scale
    ) %>%
    select(-scale)
}
```

## 3) Baseline FD-Pool results (core paper figure)

```{r baseline-fit}
fits_base <- set_names(map(crops_fd, ~ fit_fd_pool(panel_dataset_fd, .x)), crops_fd)

fd_pool_coefs <- map_df(names(fits_base), ~ extract_coefs(fits_base[[.x]], .x, "Baseline"))
fd_pool_ames  <- map_df(names(fits_base), ~ extract_ames_by_regime(fits_base[[.x]], .x, "Baseline"))

# Save paper outputs
write_csv(fd_pool_coefs, file.path(datpthout, paste0("FD_Pool_coeffs_", date_str, ".csv")))
write_csv(fd_pool_ames,  file.path(datpthout, paste0("FD_Pool_AMEs_",   date_str, ".csv")))
```

## 3a) Figure 4 — Pure warming sensitivity (pre vs post) + warming penalty band
```{r baseline-plot}
plot_df <- fd_pool_ames %>%
  rescale_ames_100() %>%
  dplyr::filter(term %in% c("Delta_GDD", "Delta_Precip")) %>%
  dplyr::mutate(
    term = dplyr::recode(term,
      "Delta_GDD"    = "ΔGDD (per 100 units)",
      "Delta_Precip" = "ΔPrecip (per 100 mm)"
    ),
    period = factor(period, levels = c("Before Break", "Post Break")),
    crop   = factor(crop, levels = crops_fd)
  )

ggplot2::ggplot(plot_df, ggplot2::aes(x = crop, y = estimate, fill = period)) +
  ggplot2::geom_bar(stat = "identity", position = ggplot2::position_dodge(width = 0.85), color = "black") +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = conf.low, ymax = conf.high),
    position = ggplot2::position_dodge(width = 0.85),
    width = 0.25
  ) +
  ggplot2::facet_wrap(~ term, scales = "free_y") +
  ggplot2::labs(
    title = "FD panel: regime-specific marginal effects (standardized shocks)",
    subtitle = "Effects on ΔYield per 100 ΔGDD / per 100 mm ΔPrecip (95% CI)",
    x = "Crop",
    y = "Marginal effect on ΔYield (t/ha)",
    fill = "Regime"
  ) +
  ggplot2::theme_minimal(base_size = 13) +
  ggplot2::theme(legend.position = "top")
```



## 3b) Figure 5 — Pure warming sensitivity (pre vs post) + warming penalty band
```{r fig5-warming-curves}
# Figure 5 — Pure warming sensitivity (pre vs post) + warming penalty band
# Uses fd_pool_coefs (Baseline) produced above

# ---- Helpers: robustly identify and strip interaction terms ----
is_delta_term <- function(term) {
  stringr::str_detect(term, "flag_before_after1")
}

strip_interaction <- function(term) {
  term %>%
    stringr::str_replace(":flag_before_after1", "") %>%
    stringr::str_replace("flag_before_after1:", "")
}

pick1 <- function(x) if (length(x) == 0) 0 else x[1]

climate_terms <- c("Delta_GDD", "I(Delta_GDD^2)")

# ---- Build β (pre) and δ (shift) tables (warming-only terms) ----
coef_warm <- fd_pool_coefs %>%
  dplyr::filter(model_type == "Baseline") %>%
  dplyr::mutate(
    is_delta  = is_delta_term(term),
    base_term = strip_interaction(term)
  ) %>%
  dplyr::filter(base_term %in% climate_terms) %>%
  dplyr::select(crop, base_term, is_delta, estimate, std.error)

beta_tbl <- coef_warm %>%
  dplyr::filter(!is_delta) %>%
  dplyr::transmute(crop, base_term, beta = estimate, beta_se = std.error)

delta_tbl <- coef_warm %>%
  dplyr::filter(is_delta) %>%
  dplyr::transmute(crop, base_term, delta = estimate, delta_se = std.error)

coef_all <- beta_tbl %>%
  dplyr::full_join(delta_tbl, by = c("crop", "base_term")) %>%
  dplyr::mutate(
    beta     = tidyr::replace_na(beta, 0),
    delta    = tidyr::replace_na(delta, 0),
    beta_se  = tidyr::replace_na(beta_se, 0),
    delta_se = tidyr::replace_na(delta_se, 0),
    post     = beta + delta
  )

# ---- Build curves per crop ----
make_warming_curve <- function(crop_id, max_gdd = 300, n = 301) {
  cc <- coef_all %>% dplyr::filter(crop == crop_id)
  if (nrow(cc) == 0) return(NULL)

  b1 <- pick1(cc$beta[cc$base_term == "Delta_GDD"])
  b2 <- pick1(cc$beta[cc$base_term == "I(Delta_GDD^2)"])
  p1 <- pick1(cc$post[cc$base_term == "Delta_GDD"])
  p2 <- pick1(cc$post[cc$base_term == "I(Delta_GDD^2)"])
  d1 <- pick1(cc$delta[cc$base_term == "Delta_GDD"])
  d2 <- pick1(cc$delta[cc$base_term == "I(Delta_GDD^2)"])

  d1_se <- pick1(cc$delta_se[cc$base_term == "Delta_GDD"])
  d2_se <- pick1(cc$delta_se[cc$base_term == "I(Delta_GDD^2)"])

  x <- seq(0, max_gdd, length.out = n)

  pre   <- b1 * x + b2 * x^2
  post  <- p1 * x + p2 * x^2
  added <- d1 * x + d2 * x^2

  # Penalty CI (delta-only), ignoring cov(d1,d2)
  se_added <- sqrt((x^2) * (d1_se^2) + (x^4) * (d2_se^2))
  added_lo <- added - 1.96 * se_added
  added_hi <- added + 1.96 * se_added

  tibble::tibble(
    crop = crop_id,
    Delta_GDD = x,
    pre = pre,
    post = post,
    added = added,
    added_lo = added_lo,
    added_hi = added_hi
  ) %>%
    tidyr::pivot_longer(cols = c(pre, post, added), names_to = "curve", values_to = "dY") %>%
    dplyr::mutate(
      dY_lo = if_else(curve == "added", added_lo, NA_real_),
      dY_hi = if_else(curve == "added", added_hi, NA_real_)
    )
}

warming_curves_all <- purrr::map_dfr(unique(coef_all$crop), make_warming_curve)

# ---- Plot (CI band only for penalty) ----
warming_curves_all <- warming_curves_all %>%
  dplyr::mutate(
    curve = factor(curve, levels = c("added", "post", "pre"),
                   labels = c("Warming Penalty", "Post-break", "Pre-break"))
  )

ggplot2::ggplot(warming_curves_all, ggplot2::aes(x = Delta_GDD, y = dY, colour = curve)) +
  ggplot2::geom_ribbon(
    data = subset(warming_curves_all, curve == "Warming Penalty"),
    ggplot2::aes(ymin = dY_lo, ymax = dY_hi, fill = curve, colour = NULL),
    alpha = 0.20
  ) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::facet_wrap(~ crop, scales = "free_y") +
  ggplot2::labs(
    title = "Pure warming sensitivity: pre- vs post-break climate slopes",
    x = "ΔGDD (warming shock)",
    y = "Predicted ΔYield (t/ha)",
    colour = NULL,
    fill = NULL
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "right")
```




## 4) Robustness: flooded-after, flooded-before, placebo (stress tests)

```{r robustness-fit}
rob_df <- dplyr::bind_rows(
  fd_pool_ames  %>% dplyr::mutate(model_type = "Baseline"),
  ames_fafter   %>% dplyr::mutate(model_type = "Flooded-After"),
  ames_fbefore  %>% dplyr::mutate(model_type = "Flooded-Before"),
  ames_placebo  %>% dplyr::mutate(model_type = "Placebo")
) %>%
  rescale_ames_100() %>%
  dplyr::filter(term %in% c("Delta_GDD","Delta_Precip")) %>%
  dplyr::mutate(
    term = dplyr::recode(term,
      "Delta_GDD"    = "ΔGDD (per 100 units)",
      "Delta_Precip" = "ΔPrecip (per 100 mm)"
    ),
    crop = factor(crop, levels = crops_fd),
    model_type = factor(model_type, levels = c("Baseline","Flooded-Before","Flooded-After","Placebo")),
    period = factor(period, levels = c("Before Break","Post Break","Placebo (pooled)"))
  )

ggplot2::ggplot(rob_df, ggplot2::aes(x = crop, y = estimate, fill = period)) +
  ggplot2::geom_bar(stat = "identity", position = ggplot2::position_dodge(width = 0.85), color = "black") +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = conf.low, ymax = conf.high),
    position = ggplot2::position_dodge(width = 0.85),
    width = 0.25
  ) +
  ggplot2::facet_grid(term ~ model_type, scales = "free_y") +
  ggplot2::labs(
    title = "Robustness stress tests (deliberate regime misclassification + placebo)",
    subtitle = "AMEs on ΔYield per 100 ΔGDD / per 100 mm ΔPrecip (95% CI)",
    x = "Crop",
    y = "Marginal effect on ΔYield (t/ha)",
    fill = "Regime"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## 5) Optional: Annex PDF (tables + sample definition + robustness summary)

```{r annex-pdf, eval=FALSE}
library(grid)
library(gridExtra)

out_dir <- file.path(datpthout, "Supplementary")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# Sample definition used in baseline FD (HAS_BREAK only)
panel_key <- sensitivity_tagged %>%
  distinct(ccc, ppp, break_year, break_tag) %>%
  arrange(ppp, ccc)

sample_counts <- panel_key %>%
  count(ppp, break_tag, name = "n_panels") %>%
  pivot_wider(names_from = break_tag, values_from = n_panels, values_fill = 0) %>%
  transmute(
    crop = ppp,
    panels_included = HAS_BREAK,
    panels_excluded = NO_BREAK,
    total_panels = HAS_BREAK + NO_BREAK
  ) %>%
  arrange(desc(panels_included))

included_panels <- panel_key %>%
  filter(break_tag == "HAS_BREAK") %>%
  transmute(crop = ppp, country = ccc, break_year) %>%
  arrange(crop, country)

excluded_panels <- panel_key %>%
  filter(break_tag == "NO_BREAK") %>%
  transmute(crop = ppp, country = ccc, reason = "No statistically validated break") %>%
  arrange(crop, country)

# Compact robustness table (ΔGDD & ΔPrecip only)
robust_table <- rob_df %>%
  mutate(
    Term = term,
    Model = as.character(model_type),
    Regime = as.character(period)
  ) %>%
  select(Model, Crop = crop, Regime, Term, estimate, std.error, conf.low, conf.high, p.value) %>%
  arrange(Term, Crop, Model, Regime)

df_to_pages <- function(df, title, rows_per_page = 28, digits = 4, font = 8) {
  df2 <- as.data.frame(df)
  num_cols <- sapply(df2, is.numeric)
  df2[num_cols] <- lapply(df2[num_cols], function(x) round(x, digits))

  n <- nrow(df2)
  idx <- seq(1, n, by = rows_per_page)
  for (s in idx) {
    e <- min(s + rows_per_page - 1, n)
    chunk <- df2[s:e, , drop = FALSE]
    tg <- tableGrob(
      chunk, rows = NULL,
      theme = ttheme_minimal(
        core    = list(fg_params = list(cex = font/10)),
        colhead = list(fg_params = list(cex = (font+1)/10, fontface = "bold"))
      )
    )
    grid.newpage()
    grid.text(title, x = 0.02, y = 0.98, just = c("left","top"),
              gp = gpar(fontface = "bold", fontsize = 12))
    grid.text(paste0("Rows ", s, "–", e, " of ", n),
              x = 0.98, y = 0.98, just = c("right","top"),
              gp = gpar(fontsize = 9, col = "grey35"))
    pushViewport(viewport(x = 0.5, y = 0.46, width = 0.98, height = 0.86))
    grid.draw(tg)
    popViewport()
  }
}

annex_file <- file.path(out_dir, paste0("Annex_1_Supplementary_FD_Panel_and_Robustness_", date_str, ".pdf"))
pdf(annex_file, width = 11, height = 8.5)

df_to_pages(fd_pool_coefs %>% arrange(crop, term),
            "Annex — FD-Pool coefficients (Baseline)", rows_per_page = 22, digits = 5)

df_to_pages(sample_counts,
            "Annex — Sample definition (counts by crop)", rows_per_page = 35, digits = 0)

df_to_pages(included_panels,
            "Annex — Included panels (HAS_BREAK)", rows_per_page = 30, digits = 0)

df_to_pages(excluded_panels,
            "Annex — Excluded panels (NO_BREAK)", rows_per_page = 30, digits = 0)

df_to_pages(robust_table,
            "Annex — Robustness summary (AMEs, standardized shocks)", rows_per_page = 22, digits = 5)

dev.off()
cat("Wrote: ", annex_file, "\n")
```

