---
title: "Supplementary Information"
format:
  docx:
    toc: true
    number-sections: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# ============================================================
# SUPPLEMENTARY INFORMATION
# Structural breaks reveal accelerating climate sensitivity
# of global cereal production
# ============================================================

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)
library(readxl)
library(knitr)
library(flextable)

datpthout <- "C:/Users/arsovsp/Documents/Climate Risk/output/"
date_str  <- format(Sys.Date(), "%Y%m%d")

paper_title <- "Structural breaks reveal accelerating climate sensitivity of global cereal production"
authors <- c(
  "Spire Arsov*",
  "Simone Pieralli"
)
corresponding_author <- "Spire Arsov"

# ------------------------------------------------------------
# Helper functions
# ------------------------------------------------------------
latest_file <- function(pattern, dir = datpthout, recursive = TRUE) {
  ff <- list.files(dir, full.names = TRUE, recursive = recursive)
  ff <- ff[grepl(pattern, basename(ff), ignore.case = TRUE)]
  if (length(ff) == 0) return(NA_character_)
  ff[which.max(file.info(ff)$mtime)]
}

latest_match <- function(patterns, dir = datpthout, recursive = TRUE) {
  for (p in patterns) {
    hit <- latest_file(p, dir = dir, recursive = recursive)
    if (!is.na(hit)) return(hit)
  }
  NA_character_
}

resolve_fig <- function(candidates, dir = datpthout) {
  for (nm in candidates) {
    p <- file.path(dir, nm)
    if (file.exists(p)) return(p)
  }
  NA_character_
}

safe_read_csv <- function(path) {
  if (is.na(path) || is.null(path) || !file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

safe_include_graphics <- function(path, note = "Figure not available.") {
  if (is.na(path) || is.null(path) || !file.exists(path)) {
    cat(note)
  } else {
    knitr::include_graphics(path)
  }
}

make_ft <- function(df, digits = 4, fontsize_body = 9) {
  if (is.null(df) || nrow(df) == 0) return(NULL)

  df2 <- df
  num_cols <- vapply(df2, is.numeric, logical(1))
  df2[num_cols] <- lapply(df2[num_cols], function(x) round(x, digits))

  flextable(df2) %>%
    theme_booktabs() %>%
    autofit() %>%
    fontsize(size = fontsize_body, part = "all") %>%
    bold(part = "header") %>%
    align(align = "center", part = "all") %>%
    valign(valign = "center", part = "all")
}

show_table_or_note <- function(df, note = "Table not available.") {
  if (inherits(df, "flextable")) {
    df
  } else if (is.null(df) || nrow(df) == 0) {
    cat(note)
  } else {
    make_ft(df)
  }
}

pstars <- function(p) dplyr::case_when(
  is.na(p) ~ "",
  p < 0.01 ~ "***",
  p < 0.05 ~ "**",
  p < 0.10 ~ "*",
  TRUE ~ ""
)
```

# Supplementary Information

## `r paper_title`

**`r paste(authors, collapse = ", ")`**

**Corresponding author:** `r corresponding_author`

**Data availability.** All data used in this study are publicly available.

**Code availability.** Code can be made available upon request.

This Supplementary Information is intentionally curated for peer review. It contains the France-specific levels diagnostics explicitly referenced in the manuscript, together with two compact panel-regression support tables that clarify the sample definition and the estimated regime shifts. Exhaustive machine-readable outputs, long annex tables, and full pipeline dumps should be kept in a separate reproducibility archive rather than in the reviewer-facing SI.

\newpage

# 1. France levels-based diagnostic case

```{r}
# ------------------------------------------------------------
# France-specific objects and figure paths
# ------------------------------------------------------------
fig_france_barley <- resolve_fig(c(
  "France_BA_Levels_Break_Diagnostic.png",
  "France_Barley_Levels_Break_Diagnostic.png"
))

fig_france_maize <- resolve_fig(c(
  "France_MA_Levels_Break_Diagnostic.png",
  "France_Maize_Levels_Break_Diagnostic.png",
  "France_Maize_Diagnostic.png"
))

fig_france_rapeseed <- resolve_fig(c(
  "France_RP_Levels_Break_Diagnostic.png",
  "France_Rapeseed_Levels_Break_Diagnostic.png"
))

fig_france_wheat <- resolve_fig(c(
  "France_WT_Levels_Break_Diagnostic.png",
  "France_Wheat_Levels_Break_Diagnostic.png"
))

france_coef_tbl <- NULL
if (exists("fr_table_show")) {
  france_coef_tbl <- fr_table_show
} else if (exists("fr_table_raw")) {
  france_coef_tbl <- fr_table_raw %>%
    transmute(
      Crop,
      Break = ifelse(is.na(Break), "—", as.character(Break)),
      `ME(GDD) Before`  = signif(me_gdd_before, 3),
      `ME(GDD) After`   = signif(me_gdd_after, 3),
      `ΔME(GDD)`        = ifelse(is.na(d_me_gdd), "NA", paste0(signif(d_me_gdd, 3), pstars(p_d_me_gdd))),
      `ME(Prec) Before` = signif(me_prec_before, 3),
      `ME(Prec) After`  = signif(me_prec_after, 3),
      `ΔME(Prec)`       = ifelse(is.na(d_me_prec), "NA", paste0(signif(d_me_prec, 3), pstars(p_d_me_prec)))
    )
}
```

Table S1 reports the France-specific levels-based coefficients and regime contrasts that accompany the country diagnostic discussion in the main text. It gives reviewers a compact numerical companion to the crop-specific breakpoint figures.

## Table S1. France levels-based coefficients and regime contrasts

```{r}
show_table_or_note(
  france_coef_tbl,
  "France coefficient table not available in this session. Render after running the France diagnostic chunk from the break-analysis script, or export a dedicated France table object first."
)
```

Figure S1 shows the France barley diagnostic in levels. It is included to document how the selected breakpoint translates into visibly different fitted relationships across regimes for a crop discussed in the France case-study set.

## Figure S1. France barley levels-based structural-break diagnostic

```{r, fig.cap="Figure S1. France barley levels-based structural-break diagnostic.", out.width="95%"}
safe_include_graphics(fig_france_barley, "Figure S1 not found.")
```

Figure S2 presents the France maize diagnostic that complements the main-text maize discussion. It shows the selected breakpoint, the regime-specific fit, and the deterioration in post-break explanatory power noted in the manuscript.

## Figure S2. France maize levels-based structural-break diagnostic

```{r, fig.cap="Figure S2. France maize levels-based structural-break diagnostic.", out.width="95%"}
safe_include_graphics(fig_france_maize, "Figure S2 not found.")
```

Figure S3 provides the France rapeseed diagnostic. It is included because rapeseed displays one of the clearest statistically supported changes in the implied levels-based response surface in the France case.

## Figure S3. France rapeseed levels-based structural-break diagnostic

```{r, fig.cap="Figure S3. France rapeseed levels-based structural-break diagnostic.", out.width="95%"}
safe_include_graphics(fig_france_rapeseed, "Figure S3 not found.")
```

Figure S4 reports the France wheat diagnostic, which complements the main text’s discussion of pronounced instability around the early 2000s and helps reviewers see the crop-specific fit deterioration directly.

## Figure S4. France wheat levels-based structural-break diagnostic

```{r, fig.cap="Figure S4. France wheat levels-based structural-break diagnostic.", out.width="95%"}
safe_include_graphics(fig_france_wheat, "Figure S4 not found.")
```

\newpage

# 2. Compact support for the first-difference panel results

```{r}
# ------------------------------------------------------------
# Panel-result fallbacks: use objects if present, saved files otherwise
# ------------------------------------------------------------
fd_pool_coefs_obj <- if (exists("fd_pool_coefs")) {
  fd_pool_coefs
} else {
  safe_read_csv(latest_match(c(
    "^FD_Pool_coeffs_[0-9]{8}\\.csv$"
  )))
}

fd_pool_ames_obj <- if (exists("fd_pool_ames")) {
  fd_pool_ames
} else if (exists("fd_pool_marginals")) {
  fd_pool_marginals
} else {
  safe_read_csv(latest_match(c(
    "^FD_Pool_AMEs_[0-9]{8}\\.csv$",
    "^FD_Pool_marginals_[0-9]{8}\\.csv$"
  )))
}

processed_breaks_obj <- if (exists("sensitivity_tagged")) {
  sensitivity_tagged
} else {
  tmp <- safe_read_csv(latest_match(c(
    "^Processed_data_breaks_flagged_[0-9]{8}\\.csv$"
  )))
  if (!is.null(tmp)) {
    tmp %>%
      mutate(
        break_tag = if_else(break_year == 2022 | is.na(break_year), "NO_BREAK", "HAS_BREAK")
      )
  } else {
    NULL
  }
}

# ------------------------------------------------------------
# Table S2: sample definition by crop
# ------------------------------------------------------------
panel_sample_tbl <- NULL
if (!is.null(processed_breaks_obj)) {
  panel_sample_tbl <- processed_breaks_obj %>%
    mutate(
      ccc = as.character(ccc),
      ppp = as.character(ppp)
    ) %>%
    distinct(ccc, ppp, break_tag) %>%
    count(ppp, break_tag, name = "n_panels") %>%
    pivot_wider(names_from = break_tag, values_from = n_panels, values_fill = 0) %>%
    mutate(total_panels = HAS_BREAK + NO_BREAK) %>%
    arrange(desc(HAS_BREAK))
}

# ------------------------------------------------------------
# Table S3: pooled coefficient decomposition (pre, shift, post)
# ------------------------------------------------------------
panel_coef_tbl <- NULL
if (!is.null(fd_pool_coefs_obj)) {
  climate_terms <- c("Delta_GDD", "I(Delta_GDD^2)", "Delta_Precip", "I(Delta_Precip^2)")

  fd_pool_core <- fd_pool_coefs_obj %>%
    mutate(
      is_delta  = str_detect(term, "flag_before_after1"),
      base_term = str_replace(term, ":flag_before_after1", "")
    ) %>%
    filter(base_term %in% climate_terms)

  if (!("p.value" %in% names(fd_pool_core))) fd_pool_core$p.value <- NA_real_
  if (!("r2_within" %in% names(fd_pool_core))) fd_pool_core$r2_within <- NA_real_

  panel_beta <- fd_pool_core %>%
    filter(!is_delta) %>%
    transmute(
      crop,
      term = base_term,
      beta_pre = estimate,
      se_beta = std.error,
      p_beta = p.value,
      r2_within = r2_within
    )

  panel_delta <- fd_pool_core %>%
    filter(is_delta) %>%
    transmute(
      crop,
      term = base_term,
      delta = estimate,
      se_delta = std.error,
      p_delta = p.value
    )

  panel_coef_tbl <- panel_beta %>%
    full_join(panel_delta, by = c("crop", "term")) %>%
    mutate(
      beta_post = beta_pre + replace_na(delta, 0),
      se_post = sqrt(se_beta^2 + replace_na(se_delta, 0)^2)
    ) %>%
    arrange(crop, match(term, climate_terms))
}

# ------------------------------------------------------------
# Table S4: regime-specific AMEs used in the main interpretation
# ------------------------------------------------------------
panel_ame_tbl <- NULL
if (!is.null(fd_pool_ames_obj)) {
  if (!("conf.low" %in% names(fd_pool_ames_obj))) fd_pool_ames_obj$conf.low <- NA_real_
  if (!("conf.high" %in% names(fd_pool_ames_obj))) fd_pool_ames_obj$conf.high <- NA_real_
  if (!("p.value" %in% names(fd_pool_ames_obj))) fd_pool_ames_obj$p.value <- NA_real_

  panel_ame_tbl <- fd_pool_ames_obj %>%
    filter(term %in% c("Delta_GDD", "Delta_Precip")) %>%
    mutate(
      shock_scale = case_when(
        term == "Delta_GDD" ~ 100,
        term == "Delta_Precip" ~ 100,
        TRUE ~ 1
      ),
      estimate = estimate * shock_scale,
      conf.low = conf.low * shock_scale,
      conf.high = conf.high * shock_scale,
      term = recode(term,
        "Delta_GDD" = "AME of ΔGDD (per 100 units)",
        "Delta_Precip" = "AME of ΔPrecip (per 100 mm)"
      )
    ) %>%
    transmute(
      crop,
      period,
      term,
      estimate,
      conf.low,
      conf.high,
      p.value = p.value
    ) %>%
    arrange(term, crop, period)
}
```

Table S2 defines the empirical base used in the first-difference panel stage. It reports, by crop, how many country–crop panels enter the break-based pooled sample and how many remain outside it because no validated break was assigned.

## Table S2. Panel sample definition by crop

```{r}
show_table_or_note(
  panel_sample_tbl,
  "Panel sample-definition table not available. Render after creating the processed break dataset or keeping the latest Processed_data_breaks_flagged file in the output folder."
)
```

Table S3 gives a compact decomposition of the pooled first-difference coefficients into pre-break slopes, shift terms, and implied post-break slopes. This lets reviewers see directly how the reported post-break effects are constructed without reproducing the full long-form regression dump.

## Table S3. First-difference pooled coefficient decomposition

```{r}
show_table_or_note(
  panel_coef_tbl,
  "Pooled coefficient decomposition not available. Render after running the FD panel script or saving the latest FD_Pool_coeffs file."
)
```

Table S4 reports the regime-specific average marginal effects underlying the main panel interpretation, expressed at the standardized 100-unit GDD shock and 100 mm precipitation shock used in the manuscript. It provides a compact numerical companion to the main panel figure without duplicating the full robustness annex.

## Table S4. Regime-specific average marginal effects at standardized shocks

```{r}
show_table_or_note(
  panel_ame_tbl,
  "Regime-specific AME table not available. Render after running the FD panel script or saving the latest FD_Pool_AMEs file."
)
```

\newpage

# 3. Note on longer outputs

The full replication workflow also produces longer annexes and machine-readable exports, including complete breakpoint tables for all country–crop pairs, variable-level break-driver dumps, long panel-inclusion lists, and extended robustness outputs. These materials are useful for full reproducibility, but they are better supplied as a separate archive or repository rather than embedded in the reviewer-facing Supplementary Information.
